---
title: "How do Parties vote in the Austrian National Council? Part II"
date: 2026-06-08
date-modified: last-modified
author: Roland Schmidt
description: |
  A follow-up to the voting patterns post, ...

categories:
  - Austria
  - parliament
  - ParlAT

reference-location: margin
knitr:
    opts_chunk:
      code-fold: show
      collapse: true
      dev: "ragg_png"
draft: false
---

# Packages

```{r}
#| label: setup
#| code-summary: "Load packages"
library(ParlAT)
library(tidyverse)
library(furrr)
library(reactable)
library(reactablefmtr)

compact_table_theme <- reactablefmtr::fivethirtyeight(font_size = 12)
table_source <- reactablefmtr::html(
  "<span style='font-size:8pt;color:grey30;font-family:Segoe UI !important;line-height:0.5'>Source: www.parlament.gv.at. Analysis: Roland Schmidt - @zoowalk.bsky.social - https://werk.statt.codes</span>"
)

style_reactable_table <- function(table, title, subtitle = NULL) {
  styled_table <- table %>%
    reactablefmtr::add_title(
      title = reactablefmtr::html(paste0(
        "<span style='font-size:12pt;'>",
        title,
        "</span>"
      ))
    )

  if (!is.null(subtitle)) {
    styled_table <- styled_table %>%
      reactablefmtr::add_subtitle(
        subtitle = reactablefmtr::html(paste0(
          "<span style='font-size:10pt;line-height:0.5;'>",
          subtitle,
          "</span>"
        )),
        font_color = "grey30",
        font_weight = "normal",
        font_style = "italic"
      )
  }

  styled_table %>%
    reactablefmtr::add_source(source = table_source)
}
```

# Get items

## Query data

```{r}
#| label: fetch-items
#| code-summary: "Fetch National Council items and cache them"
#| cache: true
#| eval: false
items_20_28 <- seq(20, 28) %>%
  map(., \(x) get_items(legis_period = x, institution = "NR")) %>%
  list_rbind()
nrow(items_20_28)
```

```{r}
#| label: load-items-cache
#| include: false
# readr::write_rds(
#   items_20_28,
#   here::here(
#     "posts",
#     "2026-06-08-ParlAT-votingPatterns-2",
#     "_data",
#     "items_20_28.rds"
#   )
# )
items_20_28 <- readr::read_rds(here::here(
  "posts",
  "2026-06-08-ParlAT-votingPatterns-2",
  "_data",
  "items_20_28.rds"
))

data_cutoff <- {
  d <- max(items_20_28$date, na.rm = TRUE)
  sub("^0", "", format(d, "%d %B %Y"))
}
```

## Keep items of interest

```{r}
#| label: define-item-scope
#| code-summary: "Select legislative item types"
items_select <- c("^A$", "RV", "BUA", "GABR", "GABR13", "BRA", "^EBR$") %>%
  paste0(., collapse = "|")

items_scope <- items_20_28 %>%
  filter(str_detect(type_doc, regex(items_select)))

items_scope %>%
  distinct(type_doc_long, type_doc)
```

# Get item details

## Query data

```{r}
#| label: fetch-item-details
#| code-summary: "Fetch item details with stages and cache them"
#| cache: true
#| eval: false
plan(multisession, workers = 4)

item_urls <- items_scope %>%
  pull(item_url)

safe_get_item_details <- safely(get_item_details)

item_details_raw <- item_urls %>%
  future_map(
    \(x) {
      safe_result <- safe_get_item_details(item_url = x, stages = TRUE)
      safe_result$item_url <- x
      safe_result
    },
    .progress = TRUE,
    .options = furrr_options(packages = c("ParlAT", "purrr", "tibble"))
  )
```

```{r}
#| label: load-item-details-cache
#| eval: true
#| include: false
readr::write_rds(
  item_details_raw,
  here::here(
    "posts",
    "2026-06-08-ParlAT-votingPatterns-2",
    "_data",
    "item_details_raw.rds"
  )
)

plan(sequential)

item_details_raw <- readr::read_rds(here::here(
  "posts",
  "2026-06-08-ParlAT-votingPatterns-2",
  "_data",
  "item_details_raw.rds"
))
```

## Prepare stages data

- check for failed extractions
- combine successful extractions into single dataframe
- combined dataframe still contains all rows for each item (also rows without a vote)

```{r}
#| label: prepare-stages-data
#| code-summary: "Unnest stages, filter vote rows, add reading and vote_text_list columns"
failed_items <- item_details_raw |>
  keep(\(x) !is.null(x$error)) |>
  map(\(x) tibble(item_url = x$item_url, error = conditionMessage(x$error))) |>
  list_rbind()

item_details <- item_details_raw |>
  keep(\(x) is.null(x$error)) |>
  map("result") |>
  list_rbind()

# unnest stages into one row per stage per item
item_stages <- item_details |>
  select(
    item_url,
    legis_period,
    type_doc,
    title,
    item_number,
    status_number,
    status_description,
    stages,
    votes
  ) |>
  filter(map_lgl(stages, \(x) !is.null(x))) |>
  unnest(stages) %>%
  mutate(stage_date = lubridate::dmy(stage_date))


data_cutoff <- max(item_stages$stage_date)
```

# Get items in scope

Categorize each extracted row.

\- only NR

\- only where votes were casted

\- which reading

\- how are votes reported

```{r}
items_scope <- item_stages |>
  filter(str_detect(phase, regex("NR"))) %>% #only NR
  filter(str_detect(stage_name, "(dritter?|zweiter?) Lesung")) %>% #only 2nd and 3rd reading
  mutate(
    reading = case_when(
      str_detect(stage_name, regex("zweite")) &
        str_detect(stage_name, regex("dritte")) ~ "second & third",
      str_detect(stage_name, regex("zweite")) ~ "second",
      str_detect(stage_name, regex("dritte")) ~ "third",
      .default = NULL
    )
  ) %>%
  filter(str_detect(
    stage_name,
    regex(
      "dafür|wechselnde Mehrheit|getrennte Abstimmung|namentliche Abstimmung|abgelehnt|mehrstimmig",
      ignore_case = T
    )
  )) %>%
  mutate(
    vote_report_type = case_when(
      str_detect(stage_name, regex("dafür", ignore_case = T)) &
        str_detect(stage_name, regex("dagegen", ignore_case = T)) ~ "regular",
      str_detect(
        stage_name,
        regex("wechselnde Mehrheit", ignore_case = T)
      ) ~ "wechselnde Mehrheit",
      str_detect(
        stage_name,
        regex("namentliche Abstimmung", ignore_case = T)
      ) ~ "namentliche Abstimmung",
      str_detect(
        stage_name,
        regex("getrennte Abstimmung", ignore_case = T)
      ) ~ "getrennte Abstimmung",
      str_detect(stage_name, regex("abgelehnt", ignore_case = T)) ~ "abgelehnt",
      str_detect(
        stage_name,
        regex("mehrstimmig", ignore_case = T)
      ) ~ "mehrstimmig",
      .default = NA
    )
  )
#sequence of vote_reporting_type clauses important. Keep regular at top.

nrow(items_scope)
nrow(item_stages)

items_scope %>%
  count(vote_report_type, sort = T) %>%
  mutate(rel = n / sum(n) * 100)
```

## Check pairs

```{r}
non_pairs <- items_scope %>%
  count(legis_period, item_url, name = "n_obs") %>%
  filter(n_obs != 2)
n_distinct(non_pairs$item_url)
```

We would expect that every time has by now only two rows, i.e., one for the second and one for the third reading.
However, there are `r n_distinct(non_pairs$item_url)` out of `r n_distinct(items_scope$item_url)` which feature more than two rows.

```{r}
non_pairs_details <- items_scope %>%
  semi_join(non_pairs, by = c("legis_period", "item_url")) %>%
  distinct() %>%
  select(
    legis_period,
    item_url,
    stage_date,
    stage_name,
    vote_report_type
  ) %>%
  mutate(n_obs = n(), .by = c("legis_period", "item_url"))

non_pairs_details_display <- non_pairs_details %>%
  arrange(desc(n_obs), legis_period, item_url, stage_date) %>%
  mutate(
    item_ref = item_url %>%
      str_remove("^https://www\\.parlament\\.gv\\.at/gegenstand/") %>%
      str_remove("\\?.*$")
  ) %>%
  select(
    legis_period,
    item_ref,
    stage_date,
    stage_name,
    vote_report_type,
    n_obs,
    item_url
  )

non_pairs_details_table <- reactable(
  non_pairs_details_display,
  columns = list(
    legis_period = colDef(
      name = "Legis. period",
      align = "right",
      minWidth = 90,
      filterable = TRUE
    ),
    item_ref = colDef(
      name = "Item",
      minWidth = 125,
      filterable = TRUE,
      cell = function(value, index) {
        htmltools::tags$a(
          href = non_pairs_details_display$item_url[index],
          target = "_blank",
          value
        )
      }
    ),
    stage_date = colDef(
      name = "Stage date",
      align = "left",
      minWidth = 110,
      filterable = TRUE
    ),
    stage_name = colDef(
      name = "Stage",
      minWidth = 540,
      filterable = TRUE,
      style = list(whiteSpace = "normal", lineHeight = "1.35")
    ),
    vote_report_type = colDef(
      name = "Vote report type",
      minWidth = 155,
      filterable = TRUE
    ),
    n_obs = colDef(
      name = "Rows per item",
      align = "right",
      minWidth = 130,
      format = colFormat(digits = 0)
    ),
    item_url = colDef(show = FALSE)
  ),
  fullWidth = TRUE,
  compact = TRUE,
  highlight = FALSE,
  outlined = TRUE,
  searchable = TRUE,
  defaultPageSize = 10,
  defaultSorted = "n_obs",
  defaultSortOrder = "desc",
  theme = compact_table_theme
) %>%
  style_reactable_table(
    "ITEMS WITH NON-STANDARD VOTE-STAGE PAIRS",
    paste0(
      "Items with more or fewer than the expected two vote-stage rows. Data as of ",
      data_cutoff,
      "."
    )
  )

non_pairs_details_table
```

As it turns out, some are duplicates in the data source, e.g. here: https://www.parlament.gv.at/gegenstand/XX/I/1530?selectedStage=105. The votes in the second and third reading are reported twice. Others are reported mulitple times because an item was within one single reading party approved/not approved and hence got two rows in the data source, i.e. https://www.parlament.gv.at/gegenstand/XXVIII/I/403

```{r}
items_scope_pairs <- items_scope %>%
  anti_join(., non_pairs, by = c("legis_period", "item_url"))
nrow(items_scope_pairs)

#check if every item has only two rows
items_scope_pairs %>%
  mutate(n_obs = n(), .by = c("legis_period", "item_url")) %>%
  filter(n_obs != 2) %>%
  nrow()

table(items_scope_pairs$reading) #TODO

items_scope_pairs_regular <-  items_scope_pairs %>%
  filter(vote_report_type=="regular")

nrow(items_scope_pairs)
nrow(items_scope_pairs_regular)
nrow(items_scope_pairs_regular)/nrow(items_scope)
```


# Extract votes

```{r}
#define function to extract reported votes
#only extract votes from vote_report_type "regular"

fn_get_votes <- function(vt) {
  vt_norm <- str_replace(
    vt,
    regex("\\.\\s*Dagegen:", ignore_case = TRUE),
    ", dagegen:"
  )

  parts <- str_split(
    vt_norm,
    regex(",\\s*dagegen:", ignore_case = TRUE),
    n = 2
  )[[1]]

  if (length(parts) != 2) {
    stop("Regular vote text does not contain both Dafür and Dagegen parts.")
  }

  if (!str_detect(parts[1], regex("Dafür:", ignore_case = TRUE))) {
    stop("Regular vote text does not contain a Dafür part.")
  }

  yes_raw <- parts[1] |>
    str_remove(regex("^.*Dafür:\\s*", dotall = TRUE, ignore_case = TRUE))

  no_raw <- parts[2]
  vote_no <- if (str_trim(no_raw) == "-") {
    character(0)
  } else {
    split_vote_parties(no_raw)
  }

  list(
    vote_yes = split_vote_parties(yes_raw),
    vote_no = vote_no
  )
}
```





```{r}
# fn to extract votes (which party voted yes/no)
valid_parties <- c(
  "SPÖ",
  "ÖVP",
  "F",
  "FPÖ",
  "GRÜNE",
  "L",
  "BZÖ",
  "STRONACH",
  "NEOS",
  "NEOS-LIF",
  "F-BZÖ",
  "JETZT",
  "PILZ",
  "NEOS/NEOS-LIF",
  "JETZT/PILZ",
  "S",
  "V",
  "N",
  "G"
)

extract_vote_tokens <- function(x) {
  if (is.na(x) || str_trim(x) == "-") {
    return(character(0))
  }

  x |>
    str_split(",\\s*") |>
    unlist() |>
    str_trim() |>
    str_remove("\\)+$") |>
    str_trim() |>
    discard(\(p) p %in% c("", "-"))
}

is_party_token <- function(x) {
  str_detect(x, regex("^[[:upper:]ÄÖÜ]+(?:[-/][[:upper:]ÄÖÜ]+)*$"))
}

extract_vote_token_candidates <- function(vt) {
  vt_norm <- str_replace(
    vt,
    regex("\\.\\s*Dagegen:", ignore_case = TRUE),
    ", dagegen:"
  )
  parts <- str_split(
    vt_norm,
    regex(",\\s*dagegen:", ignore_case = TRUE),
    n = 2
  )[[1]]
  dafuer_raw <- parts[1] |>
    str_remove(regex("^.*Dafür:\\s*", dotall = TRUE, ignore_case = TRUE))
  dafuer_parts <- str_split(
    dafuer_raw,
    regex("\\bbzw\\.?", ignore_case = TRUE)
  )[[1]]
  raw_parts <- if (length(parts) == 2) {
    c(dafuer_parts, parts[2])
  } else {
    dafuer_parts
  }

  reduce(map(raw_parts, extract_vote_tokens), union)
}

split_vote_parties <- function(x) {
  extract_vote_tokens(x) |>
    keep(is_party_token) |>
    keep(\(p) p %in% valid_parties)
}

fn_get_votes <- function(vt) {
  vt_norm <- str_replace(
    vt,
    regex("\\.\\s*Dagegen:", ignore_case = TRUE),
    ", dagegen:"
  )
  parts <- str_split(
    vt_norm,
    regex(",\\s*dagegen:", ignore_case = TRUE),
    n = 2
  )[[1]]
  dafuer_raw <- parts[1] |>
    str_remove(regex("^.*Dafür:\\s*", dotall = TRUE, ignore_case = TRUE))

  has_dagegen <- length(parts) == 2
  has_bzw <- str_detect(dafuer_raw, regex("\\bbzw\\.?", ignore_case = TRUE))

  dafuer_parts <- if (has_bzw) {
    str_split(dafuer_raw, regex("\\bbzw\\.?", ignore_case = TRUE))[[1]]
  } else {
    list(dafuer_raw)
  }

  list(
    vote_yes = reduce(map(dafuer_parts, split_vote_parties), union),
    vote_no = if (has_dagegen) split_vote_parties(parts[2]) else character(0)
  )
}

#extract votes (only for parseable report types)
items_scope <- items_scope %>%
  filter(!vote_name) %>% #remove votes with name-calls
  mutate(
    votes_li = map2(stage_name, report_type, \(vt, rt) {
      if (rt %in% c("regular", "no_dagegen")) {
        fn_get_votes(vt)
      } else {
        list(vote_yes = character(0), vote_no = character(0))
      }
    })
  )
```

## Vote extraction check

```{r}
vote_extraction_check <- items_scope %>%
  filter_out(str_detect(stage_name, regex("wechselnde"))) %>%
  mutate(vote_parties = map(votes_li, \(v) c(v$vote_yes, v$vote_no))) %>%
  select(item_url, legis_period, reading, stage_name, vote_parties) %>%
  unnest(vote_parties) %>%
  mutate(
    party_clean = str_trim(str_extract(vote_parties, "^[^(]+"))
  ) %>%
  filter(!party_clean %in% valid_parties)

vote_uppercase_unknown_check <- items_scope %>%
  mutate(vote_tokens = map(stage_name, extract_vote_token_candidates)) %>%
  select(item_url, legis_period, reading, stage_name, vote_tokens) %>%
  unnest(vote_tokens) %>%
  distinct() %>%
  filter(
    !vote_tokens %in% valid_parties,
    str_detect(vote_tokens, regex("^[[:upper:]ÄÖÜ][[:upper:]ÄÖÜ0-9/-]*$"))
  )

n_distinct(vote_extraction_check$item_url)
n_distinct(items_scope$item_url)

report_type_check <- items_scope %>%
  count(report_type)

items_scope <- items_scope %>%
  distinct()

items_scope <- items_scope %>%
  group_by(legis_period, item_url) %>%
  mutate(obs_n = n()) %>%
  ungroup()

deviant <- items_scope %>%
  filter(obs_n != 2) #limit analysis to complete pairs

item_votes_long <- items_scope %>%
  filter(report_type %in% c("regular", "no_dagegen")) %>%
  group_by(legis_period, item_url) %>%
  mutate(obs_n = n()) %>%
  ungroup() %>%
  filter(obs_n == 2) %>% # limit analysis to complete pairs
  mutate(
    votes_party = map(votes_li, \(v) {
      yes_parties <- v$vote_yes
      no_parties <- v$vote_no

      tibble(
        party = union(yes_parties, no_parties),
        yes = as.integer(party %in% yes_parties),
        no = as.integer(party %in% no_parties)
      )
    })
  ) %>%
  unnest(votes_party)

item_votes_reading_wide <- item_votes_long %>%
  filter(reading %in% c("second", "third")) %>%
  mutate(
    votes_li_raw = map_chr(votes_li, \(x) {
      paste(capture.output(str(x)), collapse = "\n")
    })
  ) %>%
  group_by(
    item_url,
    legis_period,
    type_doc,
    title,
    item_number,
    status_number,
    status_description,
    party
  ) %>%
  mutate(
    stage_name = paste(unique(stage_name), collapse = "\n\n"),
    votes_li_raw = paste(unique(votes_li_raw), collapse = "\n\n")
  ) %>%
  ungroup() %>%
  select(
    item_url,
    legis_period,
    type_doc,
    title,
    item_number,
    status_number,
    status_description,
    party,
    stage_name,
    votes_li_raw,
    reading,
    yes,
    no,
    report_type
  ) %>%
  pivot_wider(
    id_cols = c(
      item_url,
      legis_period,
      type_doc,
      title,
      item_number,
      status_number,
      status_description,
      party,
      stage_name,
      votes_li_raw
    ),
    names_from = reading,
    values_from = c(yes, no, report_type),
    names_glue = "{.value}_{reading}",
    values_fill = list(yes = NA, no = NA, report_type = NA_character_)
  ) %>%
  mutate(
    flip_pos = ifelse(no_second == 1 & yes_third == 1, TRUE, FALSE),
    flip_neg = ifelse(yes_second == 1 & no_third == 1, TRUE, FALSE)
  )
```
