---
title: "How do Parties vote in the Austrian National Council? Part II"
date: 2026-06-08
date-modified: last-modified
author: Roland Schmidt
description: |
  A follow-up to the voting patterns post, ...

categories:
  - Austria
  - parliament
  - ParlAT

reference-location: margin
knitr:
    opts_chunk:
      code-fold: show
      collapse: true
      dev: "ragg_png"
draft: false
---

# Packages

```{r}
#| label: setup
#| code-summary: "Load packages"
library(ParlAT)
library(tidyverse)
library(furrr)
```

# Get items

## Query data

```{r}
#| label: fetch-items
#| code-summary: "Fetch National Council items and cache them"
#| cache: true
#| eval: false
items_20_28 <- seq(20, 28) %>%
  map(., \(x) get_items(legis_period = x, institution = "NR")) %>%
  list_rbind()
nrow(items_20_28)
```

```{r}
#| label: load-items-cache
#| include: false
# readr::write_rds(
#   items_20_28,
#   here::here(
#     "posts",
#     "2026-06-08-ParlAT-votingPatterns-2",
#     "_data",
#     "items_20_28.rds"
#   )
# )
items_20_28 <- readr::read_rds(here::here(
  "posts",
  "2026-06-08-ParlAT-votingPatterns-2",
  "_data",
  "items_20_28.rds"
))

data_cutoff <- {
  d <- max(items_20_28$date, na.rm = TRUE)
  sub("^0", "", format(d, "%d %B %Y"))
}
```

## Keep items of interest

```{r}
#| label: define-item-scope
#| code-summary: "Select legislative item types"
items_select <- c("^A$", "RV", "BUA", "GABR", "GABR13", "BRA", "^EBR$") %>%
  paste0(., collapse = "|")

items_scope <- items_20_28 %>%
  filter(str_detect(type_doc, regex(items_select)))

items_scope %>%
  distinct(type_doc_long, type_doc)
```

# Get item details

## Query data

```{r}
#| label: fetch-item-details
#| code-summary: "Fetch item details with stages and cache them"
#| cache: true
#| eval: false
plan(multisession, workers = 4)

item_urls <- items_scope %>%
  pull(item_url)

safe_get_item_details <- safely(get_item_details)

item_details_raw <- item_urls %>%
  future_map(
    \(x) {
      safe_result <- safe_get_item_details(item_url = x, stages = TRUE)
      safe_result$item_url <- x
      safe_result
    },
    .progress = TRUE,
    .options = furrr_options(packages = c("ParlAT", "purrr", "tibble"))
  )
```

```{r}
#| label: load-item-details-cache
#| eval: true
#| include: false
# readr::write_rds(
#   item_details_raw,
#   here::here(
#     "posts",
#     "2026-06-08-ParlAT-votingPatterns-2",
#     "_data",
#     "item_details_raw.rds"
#   )
# )

# plan(sequential)

item_details_raw <- readr::read_rds(here::here(
  "posts",
  "2026-06-08-ParlAT-votingPatterns-2",
  "_data",
  "item_details_raw.rds"
))
```

## Prepare stages data

```{r}
#| label: prepare-stages-data
#| code-summary: "Unnest stages, filter vote rows, add reading and vote_text_list columns"
failed_items <- item_details_raw |>
  keep(\(x) !is.null(x$error)) |>
  map(\(x) tibble(item_url = x$item_url, error = conditionMessage(x$error))) |>
  list_rbind()

item_details <- item_details_raw |>
  keep(\(x) is.null(x$error)) |>
  map("result") |>
  list_rbind()

# unnest stages into one row per stage per item
item_stages <- item_details |>
  select(
    item_url,
    legis_period,
    type_doc,
    title,
    item_number,
    status_number,
    status_description,
    stages,
    votes
  ) |>
  filter(map_lgl(stages, \(x) !is.null(x))) |>
  unnest(stages)

# keep only rows that contain vote results, then parse them
item_stages_votes <- item_stages |>
  #filter(item_url == "https://www.parlament.gv.at/gegenstand/XX/A/1145") %>%
  filter(str_detect(phase, regex("NR"))) %>%
  filter(str_detect(
    stage_name,
    regex("Dafür|(wechselnden? Mehrheiten)", ignore_case = T)
  )) %>%
  filter(str_detect(stage_name, "(dritter?|zweiter?) Lesung")) %>%
  mutate(
    reading = case_when(
      str_detect(stage_name, regex("zweite")) &
        str_detect(stage_name, regex("dritte")) ~ "second & third",
      str_detect(stage_name, regex("zweite")) ~ "second",
      str_detect(stage_name, regex("dritte")) ~ "third",
      .default = NULL
    )
  ) %>%
  mutate(vote_name = map_lgl(
    votes,
    \(v) {
      if (is.null(v) || !"text" %in% names(v)) return(FALSE)
      any(str_detect(v$text, regex("namentlich", ignore_case = TRUE)), na.rm = TRUE)
    }
  ))

table(item_stages_votes$vote_name, item_stages_votes$reading)

item_stages_votes %>%
  filter(!vote_name) %>%
  count(reading)

# more third readings than second readings ?
item_stages_votes %>%
  count(reading)

x <- item_stages_votes %>%
  filter(!vote_name) %>%
  group_by(item_url) %>%
  summarise(
    n = n(),
    readings = list(reading)
  )
nrow(x)

y <- x |>
  filter(map_lgl(readings, \(r) is.null(r) || !"third" %in% r))

library(ParlAT)
v <- get_item_details("https://www.parlament.gv.at/gegenstand/XX/I/1581", votes=T)

# fn to extract votes

fn_get_votes <- function(vt) {
  vt_norm <- str_replace(vt, fixed(". Dagegen:"), ", dagegen:")
  parts <- str_split(vt_norm, ", dagegen:", n = 2)[[1]]
  dafuer <- parts[1] |>
    str_remove("^Dafür: ") |>
    str_split(", ") |>
    unlist() |>
    str_trim()
  gegen <- if (length(parts) == 2) {
    parts[2] |> str_split(", ") |> unlist() |> str_trim()
  } else {
    character(0)
  }
  list(dafuer = dafuer, gegen = gegen)
}

item_stages_votes <- item_stages_votes %>%
  mutate(votes = map(stage_name, \(x) fn_get_votes(x)))
```



```{r}
#| include: false
#| eval: false

item_stages_votes %>%
  count(phase)

item_stages_votes %>%
  count(stage_name)


#|>
mutate(
  stage_description = str_extract(stage_name, "^[^\n]+"),
  vote_text = str_extract(stage_name, "(?<=\n)Dafür:.+"),

  reading = case_when(
    str_detect(
      stage_description,
      regex("dritter? Lesung", ignore_case = TRUE)
    ) ~ "dritte",
    str_detect(
      stage_description,
      regex("zweiter? Lesung", ignore_case = TRUE)
    ) ~ "zweite",
    str_detect(
      stage_description,
      regex("erste[rn]? Lesung", ignore_case = TRUE)
    ) ~ "erste",
    str_detect(phase, regex("Ausschuss", ignore_case = TRUE)) ~ "Ausschuss",
    str_detect(
      stage_description,
      regex("Entschlie.*?ungsantrag", ignore_case = TRUE)
    ) ~ "Entschließungsantrag",
    str_detect(
      stage_description,
      regex("Ab.*?nderungsantrag", ignore_case = TRUE)
    ) ~ "Abänderungsantrag",
    str_detect(
      stage_description,
      regex("Antrag auf besondere Kundmachung", ignore_case = TRUE)
    ) ~ "Antrag auf bes. Kundmachung",
    str_detect(
      stage_description,
      regex("Wiederholung", ignore_case = TRUE)
    ) ~ "Wiederholung",
    str_detect(
      stage_description,
      regex("Genehmigung", ignore_case = TRUE)
    ) ~ "Genehmigung",
    str_detect(
      stage_description,
      regex("Erm.*?chtigung", ignore_case = TRUE)
    ) ~ "Ermächtigung",
    str_detect(
      stage_description,
      regex("Abstimmung", ignore_case = TRUE)
    ) ~ "Abstimmung",
    str_detect(
      stage_description,
      regex("Beschluss", ignore_case = TRUE)
    ) ~ "Beschluss",
    str_detect(
      stage_description,
      regex("Antrag", ignore_case = TRUE)
    ) ~ "Antrag",
    TRUE ~ NA_character_
  ),

```

```{r}
#| label: cache-stages-votes
#| include: false
# readr::write_rds(
#   item_stages_votes,
#   here::here(
#     "posts",
#     "2026-06-08-ParlAT-votingPatterns-2",
#     "_data",
#     "item_stages_votes.rds"
#   )
# )
item_stages_votes <- readr::read_rds(here::here(
  "posts",
  "2026-06-08-ParlAT-votingPatterns-2",
  "_data",
  "item_stages_votes.rds"
))
```

# Vote flips across readings

## Build flip dataset

```{r}
#| label: build-flips
#| code-summary: "Identify parties that flipped position between zweite and dritte Lesung"
# For bundled items with multiple votes per reading, keep the first only
readings_23 <- item_stages_votes |>
  filter(reading %in% c("zweite", "dritte")) |>
  slice_head(n = 1, by = c(item_url, reading)) |>
  filter(n_distinct(reading) == 2, .by = item_url)

# self-join to get zweite and dritte side by side per item
zweite <- readings_23 |>
  filter(reading == "zweite") |>
  select(
    item_url,
    legis_period,
    type_doc,
    title,
    item_number,
    vote_text_zweite = vote_text_list
  )

dritte <- readings_23 |>
  filter(reading == "dritte") |>
  select(item_url, vote_text_dritte = vote_text_list)

flips <- inner_join(zweite, dritte, by = "item_url") |>
  mutate(
    dafuer_zweite = map(vote_text_zweite, "dafuer"),
    gegen_zweite = map(vote_text_zweite, "gegen"),
    dafuer_dritte = map(vote_text_dritte, "dafuer"),
    gegen_dritte = map(vote_text_dritte, "gegen"),

    # gegen in zweite -> dafuer in dritte
    flip_pos = map2(gegen_zweite, dafuer_dritte, intersect),
    # dafuer in zweite -> gegen in dritte
    flip_neg = map2(dafuer_zweite, gegen_dritte, intersect),

    n_flip_pos = map_int(flip_pos, length),
    n_flip_neg = map_int(flip_neg, length)
  ) |>
  select(-vote_text_zweite, -vote_text_dritte)
```

```{r}
#| label: cache-flips
#| include: false
# readr::write_rds(
#   flips,
#   here::here(
#     "posts",
#     "2026-06-08-ParlAT-votingPatterns-2",
#     "_data",
#     "flips.rds"
#   )
# )
flips <- readr::read_rds(here::here(
  "posts",
  "2026-06-08-ParlAT-votingPatterns-2",
  "_data",
  "flips.rds"
))
```

## Visualise flips

```{r}
#| label: plot-vote-flips
#| code-summary: "Tile plot: vote flips by party and legislative period"
#| column: body-outset-right
#| fig-width: 11
#| fig-height: 6.5
#| out-width: "100%"
#| fig-align: "center"
library(ggpattern)
library(patchwork)

gov_parties_lookup <- tribble(
  ~legis_period , ~party ,
  "XX"          , "SPÖ"  , "XX"     , "ÖVP"   ,
  "XXI"         , "ÖVP"  , "XXI"    , "F"     ,
  "XXII"        , "ÖVP"  , "XXII"   , "F"     , "XXII"   , "F-BZÖ" ,
  "XXIII"       , "SPÖ"  , "XXIII"  , "ÖVP"   ,
  "XXIV"        , "SPÖ"  , "XXIV"   , "ÖVP"   ,
  "XXV"         , "SPÖ"  , "XXV"    , "ÖVP"   ,
  "XXVI"        , "ÖVP"  , "XXVI"   , "FPÖ"   ,
  "XXVII"       , "ÖVP"  , "XXVII"  , "GRÜNE" ,
  "XXVIII"      , "ÖVP"  , "XXVIII" , "SPÖ"   , "XXVIII" , "NEOS"
) |>
  mutate(is_gov = TRUE)

valid_parties <- c(
  "SPÖ",
  "ÖVP",
  "F",
  "FPÖ",
  "GRÜNE",
  "L",
  "BZÖ",
  "STRONACH",
  "NEOS",
  "NEOS-LIF",
  "F-BZÖ",
  "JETZT",
  "PILZ",
  "NEOS/NEOS-LIF",
  "JETZT/PILZ"
)

normalise_party <- function(p) {
  case_when(
    p %in% c("NEOS-LIF", "NEOS/NEOS-LIF") ~ "NEOS",
    p %in% c("JETZT", "PILZ", "JETZT/PILZ") ~ "JETZT/PILZ",
    TRUE ~ p
  )
}

# parties present in parliament per period
parties_per_period <- item_stages_votes |>
  filter(reading == "dritte") |>
  mutate(all_p = map(vote_text_list, \(x) c(x$dafuer, x$gegen))) |>
  unnest(all_p) |>
  mutate(party_clean = str_trim(str_extract(all_p, "^[^(]+"))) |>
  filter(party_clean %in% valid_parties) |>
  mutate(party = normalise_party(party_clean)) |>
  distinct(legis_period, party)

# flip long tables
pos_long <- flips |>
  filter(n_flip_pos > 0) |>
  unnest(flip_pos) |>
  rename(party = flip_pos) |>
  mutate(party = normalise_party(party)) |>
  count(legis_period, party, name = "n") |>
  mutate(direction = "gegen → dafür")

neg_long <- flips |>
  filter(n_flip_neg > 0) |>
  unnest(flip_neg) |>
  rename(party = flip_neg) |>
  mutate(party = normalise_party(party)) |>
  count(legis_period, party, name = "n") |>
  mutate(direction = "dafür → gegen")

all_parties <- union(
  union(pos_long$party, neg_long$party),
  union("ÖVP", parties_per_period$party)
)

party_order <- c(
  bind_rows(pos_long, neg_long) |>
    count(party, wt = n, sort = TRUE) |>
    pull(party),
  setdiff(
    all_parties,
    bind_rows(pos_long, neg_long) |> distinct(party) |> pull(party)
  )
)

all_periods_vec <- c(
  "XX",
  "XXI",
  "XXII",
  "XXIII",
  "XXIV",
  "XXV",
  "XXVI",
  "XXVII",
  "XXVIII"
)

make_grid <- function(long_df, dir_label) {
  expand_grid(party = all_parties, legis_period = all_periods_vec) |>
    left_join(
      long_df |> select(party, legis_period, n),
      by = c("party", "legis_period")
    ) |>
    replace_na(list(n = 0)) |>
    left_join(gov_parties_lookup, by = c("party", "legis_period")) |>
    replace_na(list(is_gov = FALSE)) |>
    left_join(
      parties_per_period |> mutate(present = TRUE),
      by = c("party", "legis_period")
    ) |>
    replace_na(list(present = FALSE)) |>
    mutate(
      direction = dir_label,
      party = factor(party, levels = rev(party_order)),
      legis_period = factor(legis_period, levels = all_periods_vec),
      label = if_else(n > 0, as.character(n), ""),
      fill_val = if_else(present, as.numeric(n), NA_real_)
    )
}

grid_all <- bind_rows(
  make_grid(pos_long, "gegen → dafür"),
  make_grid(neg_long, "dafür → gegen")
) |>
  mutate(
    direction = factor(direction, levels = c("gegen → dafür", "dafür → gegen"))
  )

pal_pos <- "#d73027"
pal_neg <- "#4575b4"
n_max <- max(grid_all$n, na.rm = TRUE)

flip_base_theme <- theme_minimal(base_size = 11, base_family = "Noto Sans") +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(size = 9),
    plot.title = element_text(
      face = "bold",
      size = 14,
      hjust = 0,
      margin = margin(b = 4)
    ),
    plot.subtitle = element_text(
      size = 10,
      lineheight = 1.3,
      colour = "grey30",
      margin = margin(b = 10)
    ),
    plot.caption = element_text(
      size = 8,
      colour = "grey45",
      hjust = 0,
      margin = margin(t = 10)
    ),
    plot.title.position = "plot",
    plot.caption.position = "plot",
    legend.position = "none"
  )

make_flip_panel <- function(data, dir_label, fill_high, show_y = TRUE) {
  d <- filter(data, direction == dir_label)
  p <- d |>
    ggplot(aes(
      x = legis_period,
      y = party,
      fill = fill_val,
      pattern = is_gov
    )) +
    geom_tile_pattern(
      colour = "white",
      linewidth = 0.6,
      pattern_colour = "grey50",
      pattern_fill = "grey50",
      pattern_density = 0.08,
      pattern_spacing = 0.06,
      pattern_alpha = 0.45
    ) +
    geom_text(
      data = \(x) filter(x, n > 0),
      aes(x = legis_period, y = party, label = label),
      colour = "white",
      size = 3.2,
      fontface = "bold",
      family = "Noto Sans",
      inherit.aes = FALSE
    ) +
    scale_fill_gradient(
      low = "grey93",
      high = fill_high,
      limits = c(0, n_max),
      na.value = "white",
      guide = "none"
    ) +
    scale_pattern_manual(
      values = c("TRUE" = "stripe", "FALSE" = "none"),
      guide = "none"
    ) +
    scale_x_discrete(position = "top") +
    labs(x = NULL, y = NULL, title = dir_label) +
    flip_base_theme +
    theme(
      plot.title = element_text(
        face = "bold",
        size = 11,
        hjust = 0.5,
        margin = margin(b = 4)
      )
    )
  if (!show_y) {
    p <- p + theme(axis.text.y = element_blank())
  }
  p
}

t_pos <- make_flip_panel(grid_all, "gegen → dafür", pal_pos, show_y = TRUE)
t_neg <- make_flip_panel(grid_all, "dafür → gegen", pal_neg, show_y = FALSE)

(t_pos | t_neg) +
  plot_annotation(
    title = "Vote flips between zweite and dritte Lesung",
    subtitle = paste0(
      "Number of items where a party switched position between second and third reading. ",
      "White = not in parliament. Grey = in parliament, no flips.\n",
      "Diagonal stripe = in government. Left: gegen→dafür. Right: dafür→gegen."
    ),
    caption = "Data: www.parlament.gv.at via ParlAT. Analysis: Roland Schmidt - @zoowalk.bsky.social",
    theme = theme(
      plot.title = element_text(
        face = "bold",
        size = 14,
        hjust = 0,
        family = "Noto Sans",
        margin = margin(b = 4)
      ),
      plot.subtitle = element_text(
        size = 10,
        colour = "grey30",
        lineheight = 1.3,
        family = "Noto Sans",
        margin = margin(b = 8)
      ),
      plot.caption = element_text(
        size = 8,
        colour = "grey45",
        hjust = 0,
        family = "Noto Sans",
        margin = margin(t = 8)
      )
    )
  )
```

## Amendment voting as a legislative strategy

The stages data also contains votes on *Abänderungsanträge* (amendment motions). This lets us ask a more nuanced question: when a party votes for an amendment on a bill it previously opposed, does that amendment act as a "price of support" — converting a gegen vote in the zweite Lesung into a dafür vote in the dritte?

Crucially, we need to condition on whether the amendment actually *passed*. If the government majority rejected an opposition amendment, staying against in the dritte is completely expected — nothing changed. The interesting cases are only those where the amendment passed.

```{r}
#| label: build-amendment-pathways
#| code-summary: "Build amendment pathway analysis"
normalise_party <- function(p) {
  case_when(
    p %in% c("NEOS-LIF", "NEOS/NEOS-LIF") ~ "NEOS",
    p %in% c("JETZT", "PILZ", "JETZT/PILZ") ~ "JETZT/PILZ",
    TRUE ~ p
  )
}

gov_parties_lookup <- tribble(
  ~legis_period , ~party ,
  "XX"          , "SPÖ"  , "XX"     , "ÖVP"   ,
  "XXI"         , "ÖVP"  , "XXI"    , "F"     ,
  "XXII"        , "ÖVP"  , "XXII"   , "F"     , "XXII"   , "F-BZÖ" ,
  "XXIII"       , "SPÖ"  , "XXIII"  , "ÖVP"   ,
  "XXIV"        , "SPÖ"  , "XXIV"   , "ÖVP"   ,
  "XXV"         , "SPÖ"  , "XXV"    , "ÖVP"   ,
  "XXVI"        , "ÖVP"  , "XXVI"   , "FPÖ"   ,
  "XXVII"       , "ÖVP"  , "XXVII"  , "GRÜNE" ,
  "XXVIII"      , "ÖVP"  , "XXVIII" , "SPÖ"   , "XXVIII" , "NEOS"
)

gov_by_period <- gov_parties_lookup |>
  group_by(legis_period) |>
  summarise(gov_parties = list(party), .groups = "drop")

stages_3 <- item_stages_votes |>
  filter(reading %in% c("zweite", "Abänderungsantrag", "dritte")) |>
  select(item_url, legis_period, reading, vote_text_list)

zweite_df <- stages_3 |>
  filter(reading == "zweite") |>
  select(item_url, legis_period, vtl_zweite = vote_text_list)

dritte_df <- stages_3 |>
  filter(reading == "dritte") |>
  group_by(item_url) |>
  slice_tail(n = 1) |>
  ungroup() |>
  select(item_url, vtl_dritte = vote_text_list)

# Determine amendment outcome: passed if government parties voted dafür
amendment_df <- stages_3 |>
  filter(reading == "Abänderungsantrag") |>
  left_join(gov_by_period, by = "legis_period") |>
  mutate(
    dafuer_amend_raw = map(vote_text_list, \(x) {
      map_chr(x$dafuer, normalise_party)
    }),
    gegen_amend_raw = map(vote_text_list, \(x) {
      map_chr(x$gegen, normalise_party)
    }),
    amendment_passed = map2_lgl(gov_parties, dafuer_amend_raw, \(g, d) {
      any(g %in% d)
    }) &
      !map2_lgl(gov_parties, gegen_amend_raw, \(g, a) any(g %in% a))
  ) |>
  group_by(item_url) |>
  slice_tail(n = 1) |>
  ungroup() |>
  select(item_url, vtl_amendment = vote_text_list, amendment_passed)

amendment_pathways <- zweite_df |>
  inner_join(dritte_df, by = "item_url", relationship = "many-to-many") |>
  inner_join(amendment_df, by = "item_url", relationship = "many-to-many") |>
  distinct(item_url, .keep_all = TRUE) |>
  mutate(
    gegen_zweite = map(vtl_zweite, \(x) map_chr(x$gegen, normalise_party)),
    dafuer_dritte = map(vtl_dritte, \(x) map_chr(x$dafuer, normalise_party)),
    gegen_dritte = map(vtl_dritte, \(x) map_chr(x$gegen, normalise_party)),
    dafuer_amend = map(vtl_amendment, \(x) map_chr(x$dafuer, normalise_party)),
    gegen_amend = map(vtl_amendment, \(x) map_chr(x$gegen, normalise_party)),
    # gegen zweite → dafür amendment → dafür dritte  ("price of support")
    flip_via_amend = pmap(
      list(gegen_zweite, dafuer_amend, dafuer_dritte),
      \(gz, da, dd) intersect(intersect(gz, da), dd)
    ),
    # gegen zweite → gegen/absent amendment → dafür dritte  (flip without amendment)
    flip_no_amend = pmap(
      list(gegen_zweite, gegen_amend, dafuer_dritte),
      \(gz, ga, dd) intersect(intersect(gz, ga), dd)
    ),
    # gegen zweite → dafür amendment → gegen dritte
    amend_no_flip = pmap(
      list(gegen_zweite, dafuer_amend, gegen_dritte),
      \(gz, da, gd) intersect(intersect(gz, da), gd)
    ),
    n_flip_via_amend = map_int(flip_via_amend, length),
    n_flip_no_amend = map_int(flip_no_amend, length),
    n_amend_no_flip = map_int(amend_no_flip, length)
  )
```

Among the `r nrow(amendment_pathways)` items with all three stages, `r sum(amendment_pathways$amendment_passed)` had a *passed* amendment and `r sum(!amendment_pathways$amendment_passed)` had a *failed* amendment. Conditioning on amendment outcome changes the picture entirely:

```{r}
#| label: amendment-pathway-summary
#| code-summary: "Pathway counts by amendment outcome"
bind_rows(
  amendment_pathways |>
    filter(amendment_passed) |>
    summarise(
      `Amendment passed` = "Yes",
      `Price of support (gegen→amend✓→dritte dafür)` = sum(
        n_flip_via_amend > 0
      ),
      `Stayed gegen despite passed amendment` = sum(n_amend_no_flip > 0),
      `Flip without amendment` = sum(n_flip_no_amend > 0)
    ),
  amendment_pathways |>
    filter(!amendment_passed) |>
    summarise(
      `Amendment passed` = "No",
      `Price of support (gegen→amend✓→dritte dafür)` = sum(
        n_flip_via_amend > 0
      ),
      `Stayed gegen despite passed amendment` = sum(n_amend_no_flip > 0),
      `Flip without amendment` = sum(n_flip_no_amend > 0)
    )
) |>
  knitr::kable()
```

When the amendment *failed*, opposition parties almost always stayed against in the dritte — exactly as expected. When the amendment *passed*, 10 items show a party flipping to dafür, while only 3 items show a party that supported the (successful) amendment but still voted gegen. Those 3 cases are the genuinely puzzling ones: the amendment they demanded was incorporated, yet they did not follow through with a final yes vote.

```{r}
#| label: plot-amendment-pathways
#| code-summary: "Tile plot: amendment pathways by party and legislative period"
#| column: body-outset-right
#| fig-width: 11
#| fig-height: 5
#| out-width: "100%"
#| fig-align: "center"
all_periods_vec <- c(
  "XX",
  "XXI",
  "XXII",
  "XXIII",
  "XXIV",
  "XXV",
  "XXVI",
  "XXVII",
  "XXVIII"
)

make_pathway_grid <- function(df, col, label) {
  long <- df |>
    filter(map_int(.data[[col]], length) > 0) |>
    unnest(!!sym(col)) |>
    rename(party = !!sym(col)) |>
    count(legis_period, party, name = "n")

  all_parties_p <- union(long$party, "ÖVP")

  expand_grid(party = all_parties_p, legis_period = all_periods_vec) |>
    left_join(long, by = c("party", "legis_period")) |>
    replace_na(list(n = 0)) |>
    mutate(
      pathway = label,
      party = factor(
        party,
        levels = rev(c(
          long |> count(party, wt = n, sort = TRUE) |> pull(party),
          setdiff(all_parties_p, long |> distinct(party) |> pull(party))
        ))
      ),
      legis_period = factor(legis_period, levels = all_periods_vec),
      label_txt = if_else(n > 0, as.character(n), "")
    )
}

passed <- filter(amendment_pathways, amendment_passed)

grid_price <- make_pathway_grid(
  passed,
  "flip_via_amend",
  "Amendment passed → flip\n(gegen→amend dafür→dritte dafür)"
)
grid_strategic <- make_pathway_grid(
  passed,
  "amend_no_flip",
  "Amendment passed → still gegen\n(gegen→amend dafür→dritte gegen)"
)

pal_price <- "#d73027"
pal_strategic <- "#4575b4"

n_max_p <- max(c(grid_price$n, grid_strategic$n), na.rm = TRUE)

make_amend_panel <- function(data, fill_high, show_y = TRUE) {
  p <- data |>
    ggplot(aes(x = legis_period, y = party, fill = n)) +
    geom_tile(colour = "white", linewidth = 0.6) +
    geom_text(
      data = \(x) filter(x, n > 0),
      aes(label = label_txt),
      colour = "white",
      size = 3.2,
      fontface = "bold",
      family = "Noto Sans"
    ) +
    scale_fill_gradient(
      low = "grey93",
      high = fill_high,
      limits = c(0, n_max_p),
      guide = "none"
    ) +
    scale_x_discrete(position = "top") +
    labs(x = NULL, y = NULL, title = unique(data$pathway)) +
    theme_minimal(base_size = 11, base_family = "Noto Sans") +
    theme(
      panel.grid = element_blank(),
      axis.text = element_text(size = 9),
      plot.title = element_text(
        face = "bold",
        size = 10,
        hjust = 0.5,
        margin = margin(b = 4)
      ),
      legend.position = "none"
    )
  if (!show_y) {
    p <- p + theme(axis.text.y = element_blank())
  }
  p
}

p_price <- make_amend_panel(grid_price, pal_price, show_y = TRUE)
p_strategic <- make_amend_panel(grid_strategic, pal_strategic, show_y = FALSE)

(p_price | p_strategic) +
  plot_annotation(
    title = "Among items where the amendment passed: did opposing parties follow through?",
    subtitle = paste0(
      "Only items where the amendment was adopted (government voted dafür). ",
      "Left: party voted gegen (zweite), for the amendment, then dafür (dritte) — amendment as price of support.\n",
      "Right: party voted gegen (zweite), for the amendment, yet stayed gegen (dritte) — ",
      "amendment insufficient to secure final support."
    ),
    caption = "Data: www.parlament.gv.at via ParlAT. Analysis: Roland Schmidt - @zoowalk.bsky.social",
    theme = theme(
      plot.title = element_text(
        face = "bold",
        size = 13,
        hjust = 0,
        family = "Noto Sans",
        margin = margin(b = 4)
      ),
      plot.subtitle = element_text(
        size = 9,
        colour = "grey30",
        lineheight = 1.3,
        family = "Noto Sans",
        margin = margin(b = 8)
      ),
      plot.caption = element_text(
        size = 8,
        colour = "grey45",
        hjust = 0,
        family = "Noto Sans",
        margin = margin(t = 8)
      )
    )
  )
```