---
title: "Introducing {ParlAT}: An R Package to access Open Data of the Austrian Parliament"
date: 2026-02-14
draft: true
date-modified: last-modified
author: Roland Schmidt
description: |
  {ParlAT} is an R package that wraps the Austrian Parliament's Open Data API —
  making it straightforward to query data on MPs, legislative items, committees, plenary meetings, and more from R.

categories:
  - Austria
  - parliament
  - r-bloggers

reference-location: margin
knitr:
    opts_chunk:
      code-fold: show
      collapse: true
      dev: "ragg_png"
dfraft: true
---

```{r}
#| label: setup
#| include: false
#| code-summary: "Setup"
library(ParlAT)
library(tidyverse)
```

Anyone who has followed this blog will know that I frequently rely on data published by the Austrian Parliament — whether it's for analyzing [speeches](posts/2024-06-11-parliament-speech-length-limits/), tracing [calls to order](posts/2024-09-09-call-for-order/), or examining [the prior experience of MPs](posts/2024-10-24-mps-experience/). Until now, most of that work required custom scripts to query the Parliament's API and wrangle the JSON responses into tidy data frames. The [{ParlAT}](https://werkstattcodes.github.io/ParlAT/) package is an attempt to bundle these steps into a consistent, reusable interface. Overall, though, the package was developed with the intent to make the data of the Parliament more accessible to anyone interested in the Parliament's affairs, be it researchers, students, journalists etc, and working in R.

At this point, special thanks to Simon Hofer, Open Data evangelist at the Austrian Parliament, who was super helpful when diving into the details of the API. However, please note that neither the package nor its author is affiliated with the Austrian Parliament.

The package is still in development, but most of the functionality (and quite a bit more) is in place. Below I showcase **only some** examples.

You can install {ParlAT} from GitHub:

``` r
# install.packages("pak")
pak::pak("werkstattcodes/ParlAT")
```

For the full function reference and a more detailed introductory vignette, see the [documentation site](https://werkstattcodes.github.io/ParlAT/).

## What data is available?

The Austrian Parliament publishes [25 open datasets](https://www.parlament.gv.at/recherchieren/open-data/daten-und-lizenz/index.html) — with some exceptions — under a CC BY 4.0 license. These cover a broad range of parliamentary activity: MP records going back to 1918, committee compositions and memberships, plenary meeting protocols, citizen initiatives (Bürgerinitiativen), government bills (Regierungsvorlagen), written questions (Schriftliche Anfragen), EU submissions, and more.

{ParlAT} mirrors the API's structure. Most datasets map directly to a dedicated function; some, like the many types of parliamentary items (Verhandlungsgegenstände), are accessed through `get_items()` with different `item` parameters. Here's a condensed overview of the main functions:

| Function | Data |
|-----------------------------|-------------------------------------------|
| `get_mps()` | MPs since 1918, by period or date |
| `get_mps_current()` | Current composition of NR or BR |
| `get_mps_details()` | Individual MP speeches, activities, committee work |
| `get_mandates()` | All mandates of one or more individuals |
| `get_items()` | Parliamentary items: bills, motions, questions, etc. |
| `get_committees()` | Committees, optionally with member lists |
| `get_plenary_meetings()` | Plenary meeting metadata |
| `get_transcripts()` | Stenographic protocols with PDF export |
| `get_events()` | Parliamentary events and calendar |
| `get_participation()` | Public participation / consultation data |

## MPs: Current composition

A natural starting point is the composition of Parliament itself. `get_mps_current()` retrieves the members currently serving in the National Council (Nationalrat) or the Federal Council (Bundesrat):

```{r}
#| eval: true
#| code-summary: "Current NR composition"
df_current <- get_mps_current(institution = "NR")

glimpse(df_current)

df_current |>
  count(party, sort = TRUE)
```

The function also works for the Federal Council (Bundesrat):

```{r}
#| eval: true
#| code-summary: "Current BR composition"
df_current_br <- get_mps_current(institution = "BR")
nrow(df_current_br)
```

## MPs: Historical queries

For historical queries, `get_mps()` returns MPs active during a given legislative period or on a specific date. The function accepts both numeric and Roman numeral period identifiers (e.g., `27` or `"XXVII"`), as well as special codes like `"PN"` (Provisorische Nationalversammlung) and `"KN"` (Konstituierende Nationalversammlung).

```{r}
#| code-summary: "Composition of the National Council per date"
mpsEUMembership <- get_mps(date = "10.06.1994", institution = "NR")
nrow(mpsEUMembership)
glimpse(mpsEUMembership)
```

Unnesting the details column provides further info.

```{r}
#| code-summary: "Unnesting details"
mpsEUMembershipDetails <- mpsEUMembership %>%
  unnest(mp_details)

glimpse(mpsEUMembershipDetails)
```

### By legislative period

```{r}
#| eval: true
#| code-summary: "MPs by legislative period"
# All MPs of the National Council during the 22nd legislative period
mps_legis22 <- get_mps(legis_period = 22, institution = "NR")
```

Each row corresponds to one MP per legislative period. A nested `mp_details` column captures mandate-level details — party affiliation, electoral list, position — which can change within a single period. This is where things get interesting: unnesting `mp_details` reveals party switches, changes in parliamentary group, or mandate interruptions within a period.

### By date

When a specific date is more appropriate than a legislative period — for instance, when working with the Federal Council, which doesn't follow legislative periods in the same way — `get_mps()` accepts a `date` parameter:

```{r}
#| eval: true
#| code-summary: "MPs by date (BR)"
# Members of the Federal Council on 30 October 2013
df_mps_br <- get_mps(date = "30.10.2013", institution = "BR")
nrow(df_mps_br)
```

Next to the data retrieval functions, there are some auxiliary functions which hopefully come handy. E.g. to find the relevant dates for a specific legislative period, `get_legis_periods()` provides the start and end dates. If i have a date, it returns the pertaining legislative period.

```{r}
#| eval: true
#| code-summary: "Legislative period dates"
get_legis_periods(legis_period = 25)

get_legis_periods(date = "29.03.1977")
```

### Filtering by gender, party, state

`get_mps()` supports a number of additional filters. These can be combined:

```{r}
#| eval: true
#| code-summary: "Filtering MPs by gender, party, state"
# Female SPÖ MPs in the current legislative period
female_spoe <- get_mps(
  gender = "female",
  party = "SPÖ",
  legis_period = 28,
  institution = "NR"
)
female_spoe

# MPs from Vienna during the 27th period
vienna_mps <- get_mps(
  state = "Wien",
  legis_period = 27,
  institution = "NR"
)
vienna_mps
```

## Mandates: Tracing careers across Parliament

`get_mandates()` returns all current and past mandates for one or more individuals. It accepts either names or unique `pad_intern` identifiers. Mandates are not limited to Parliament — they also cover executive positions (e.g., Chancellor, Minister), presidencies, and other functions.

```{r}
#| eval: true
#| code-summary: "Get mandates by name"
get_mandates(name = c("Karl Nehammer", "Andreas Babler"))
```

The returned data includes `position_code` (e.g., "NR", "BK" for Bundeskanzler, "BM" for Bundesminister), start and end dates, party affiliation, and a flag for whether the mandate is currently active. This makes it straightforward to calculate the total time an MP has served:

```{r}
#| eval: true
#| code-summary: "Service duration calculation"
# How long have current NR members served in the National Council?
df_mandates <- get_mandates(
  pad_intern = df_current$pad_intern,
  institution = "NR"
)

duration <- df_mandates |>
  # For active mandates, use today as end date
  mutate(
    position_date_end = case_when(
      is.na(position_date_end) & position_active == TRUE ~ today(),
      .default = position_date_end
    )
  ) |>
  mutate(
    days = as.numeric(difftime(
      position_date_end,
      position_date_start,
      units = "days"
    ))
  ) |>
  summarise(total_days = sum(days), .by = c(pad_intern, name)) |>
  arrange(desc(total_days))

# Top 5 longest-serving current NR members
duration |> slice_head(n = 5)
```

`get_mandates()` also accepts a `date` argument to filter for mandates active at a specific point in time, and an `institution` argument to restrict to a single chamber.

## Identifying MPs: `get_pad_intern()` and `get_names()`

Two further examples of auxiliary functions are `get_pat`

The unique identifier for each person in the Parliament's database is `pad_intern`. Many functions accept it as input. To look it up by name:

```{r}
#| eval: true
#| code-summary: "Look up pad_intern ID"
get_pad_intern("Strache")
```

`get_names()` retrieves all name variants an MP has used over time — relevant when names change due to marriage, divorce, or other reasons. A good example is Pia Philippa Beck (formerly Strache):

```{r}
#| eval: true
#| code-summary: "Name variants over time"
# All name variants

get_names(pad_intern = 44127)
```

## Parliamentary items (Verhandlungsgegenstände)

`get_items()` is arguably the most versatile function in the package. It covers a broad spectrum of parliamentary business — motions (Anträge), committee reports (Ausschussberichte), government bills (Regierungsvorlagen), written questions (Anfragen), EU submissions, citizen initiatives (Bürgerinitiativen), popular initiatives (Volksbegehren), and more — all controlled through the `item` parameter.

### Government bills

```{r}
#| eval: true
#| code-summary: "Government bills"
# Government bills tabled in the 27th legislative period
gov_bills <- get_items(item = "RV", legis_period = 27)

glimpse(gov_bills)
```

### Written questions across multiple periods

```{r}
#| eval: true
#| code-summary: "Written questions across periods"
# Written and oral questions from the 20th to 27th period
questions <- get_items(item = "J_JPR_M", legis_period = seq(20, 27))

nrow(questions)
```

The returned data includes metadata such as topic classification (`topics` is a list-column), sponsoring parliamentary group(s), and document type. This makes it straightforward to aggregate and visualise parliamentary activity — for example, counting questions by party and legislative period.

```{r}
#| label: fig-questions-by-party
#| fig-cap: "Written and oral questions (J_JPR_M) filed per parliamentary group across legislative periods 20–27, National Council."
#| eval: true
#| code-summary: "Visualise questions by party across periods"
party_colors <- c(
  "FPÖ" = "#0056A2",
  "BZÖ" = "#F5831F",
  "GRÜNE" = "#78A22F",
  "JETZT/Pilz" = "#1C3A5E",
  "NEOS" = "#E84188",
  "ÖVP" = "#62C2CE",
  "SPÖ" = "#E2001A",
  "Stronach" = "#888888"
)

questions |>
  tidyr::unnest(parl_group) |>
  mutate(
    party = case_match(
      parl_group,
      c("F", "F-BZÖ", "FPÖ") ~ "FPÖ",
      "BZÖ" ~ "BZÖ",
      "GRÜNE" ~ "GRÜNE",
      c("JETZT", "PILZ") ~ "JETZT/Pilz",
      c("L", "NEOS-LIF", "NEOS") ~ "NEOS",
      "ÖVP" ~ "ÖVP",
      "SPÖ" ~ "SPÖ",
      "STRONACH" ~ "Stronach",
      .default = NA_character_
    ),
    period_int = as.integer(as.roman(legis_period))
  ) |>
  filter(!is.na(party), institution == "NR") |>
  count(period_int, party) |>
  ggplot(aes(x = period_int, y = n, colour = party)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  scale_colour_manual(values = party_colors, name = NULL) +
  scale_x_continuous(breaks = 20:27) +
  labs(
    x = "Legislative period",
    y = "Number of questions",
    title = "Parliamentary questions by party, periods 20–27",
    subtitle = "National Council — written and oral questions (J, JPR, M)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title.position = "plot",
    legend.position = "right",
    panel.grid.minor = element_blank()
  )
```

### Filtering by topic and parliamentary group

Items can be filtered by topic (Thema), parliamentary group, EuroVoc terms, and more. The `parl_group_names_standard` argument expands party abbreviations to their historical variants — so searching for "FPÖ" also captures items filed under "F" or "F-BZÖ":

```{r}
#| eval: true
#| code-summary: "Filter by topic and party"
# SPÖ motions on health topics during the 27th period
spoe_health <- get_items(
  item = "ANTR",
  topic = "Gesundheit und Ernährung",
  parl_group = "SPÖ",
  legis_period = 27,
  institution = "NR",
  parl_group_names_standard = TRUE
)
```

## An MP's work in detail

`get_mps_details()` provides granular data on individual MPs' plenary speeches, committee memberships, and parliamentary activities. It takes a `pad_intern` and a `detail_type` ("plenary", "committees", or "activities").

### Plenary speeches

```{r}
#| eval: true
#| code-summary: "Plenary speeches"
# All plenary speeches by Stephanie Krisper in the National Council
get_pad_intern("Krisper")

df_plenary_krisper <- get_mps_details(
  pad_intern = 2344,
  detail_type = "plenary",
  institution = "NR",
  echo = T
)

df_plenary_krisper |>
  unnest(position_name) |>
  count(position_name, sort = TRUE)
```

`get_mps_details()` returns all speeches regardless of the person's role at the time. The `position_name` list-column reveals which position(s) the person held when the speech was given.

A natural follow-up is to visualise *when* those speeches happened. The `date` column maps directly onto a GitHub-style calendar heatmap:

```{r}
#| label: fig-krisper-calendar
#| fig-cap: "GitHub-style calendar heatmap of Stephanie Krisper's plenary speeches. Each cell is one day; colour depth encodes speech count using GitHub's five-level contribution palette."
#| eval: true
#| code-summary: "Calendar heatmap of Krisper's speeches"
#| fig-height: 7
#| fig-width: 10
# GitHub's five-level contribution palette
gh_pal <- c(
  "0" = "#ebedf0",
  "1" = "#9be9a8",
  "2–3" = "#40c463",
  "4–6" = "#30a14e",
  "7+" = "#216e39"
)

# Full date spine — silent days render as lightest tile, not blank space
date_min <- lubridate::floor_date(min(df_plenary_krisper$date), "year")
date_max <- lubridate::ceiling_date(max(df_plenary_krisper$date), "year") - 1

df_cal <- tibble(date = seq.Date(date_min, date_max, by = "day")) |>
  mutate(
    iso_year = lubridate::isoyear(date), # keeps Dec/Jan ISO-week edge days in the right facet
    week_no = lubridate::isoweek(date),
    dow = lubridate::wday(date, label = TRUE, abbr = TRUE, week_start = 1)
  ) |>
  left_join(count(df_plenary_krisper, date, name = "n"), by = "date") |>
  replace_na(list(n = 0)) |>
  mutate(
    n_cat = cut(
      n,
      breaks = c(-Inf, 0, 1, 3, 6, Inf),
      labels = names(gh_pal),
      right = TRUE
    )
  )

ggplot(df_cal, aes(x = week_no, y = fct_rev(dow), fill = n_cat)) +
  geom_tile(width = 0.85, height = 0.85, colour = NA) +
  facet_wrap(~iso_year, ncol = 1, strip.position = "left") +
  scale_fill_manual(values = gh_pal, name = "Speeches", drop = FALSE) +
  scale_x_continuous(
    breaks = c(1, 5, 9, 14, 18, 22, 27, 31, 35, 40, 44, 48),
    labels = month.abb,
    expand = c(0.01, 0)
  ) +
  scale_y_discrete(
    breaks = c("Mon", "Wed", "Fri"), # GitHub only labels alternating days
    expand = c(0, 0.5)
  ) +
  coord_fixed() +
  labs(
    x = NULL,
    y = NULL,
    title = "Stephanie Krisper — plenary speeches",
    subtitle = "National Council"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title.position = "plot",
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    strip.text.y.left = element_text(angle = 0, size = 9, hjust = 1),
    legend.position = "bottom",
    legend.key.size = unit(0.55, "cm"),
    legend.direction = "horizontal",
    panel.spacing.y = unit(0.6, "cm"),
    plot.background = element_rect(fill = "white", colour = NA)
  )
```

To get a sense of what Krisper was actually speaking *about*, we can trace the exact items she debated. The `meeting_url` in the plenary data and `url_meeting` in `get_plenary_meetings()` refer to the same sessions but use different URL schemes (`/sitzung/` vs `/gegenstand/`), so a small normalisation step aligns them. That gives us `url_item` for each item in her sessions. We then call `get_item_details()` per item, which returns a `speeches` nested column listing every speaker per stage. Filtering to Krisper's name leaves only items she personally contributed to, each carrying `topics`, `headwords`, and `eurovoc`.

```{r}
#| label: fig-krisper-topics
#| fig-cap: "Topics of parliamentary items where Stephanie Krisper personally gave a speech."
#| eval: true
#| cache: true
#| code-summary: "Topics of items Krisper actually spoke on"
# Legislative periods as integers
krisper_periods <- df_plenary_krisper |>
  distinct(legis_period) |>
  mutate(p = as.integer(as.roman(legis_period))) |>
  pull(p)

# Step 1: Items on the agenda in all sessions where Krisper spoke
df_meeting_acts <- map(
  krisper_periods,
  \(p) {
    get_plenary_meetings(
      institution = "NR",
      legis_period = p,
      meeting_and_activities = "activities"
    )
  }
) |>
  list_rbind()

# meeting_url uses /sitzung/XXVI/NRSITZ/7?selectedStage=111
# url_meeting uses /gegenstand/XXVI/NRSITZ/7 — same meeting, different URL scheme
# normalise: swap path prefix and strip query string before joining
krisper_joined <- df_plenary_krisper |>
  distinct(meeting_url) |>
  mutate(
    url_meeting_norm = meeting_url |>
      str_replace("/sitzung/", "/gegenstand/") |>
      str_remove("\\?.*")
  ) |>
  left_join(df_meeting_acts, by = c("url_meeting_norm" = "url_meeting"))

unmatched_k <- krisper_joined |> filter(is.na(url_item)) |> pull(meeting_url)
if (length(unmatched_k) > 0) {
  message(sprintf(
    "%d meeting(s) unmatched in activities data:\n%s",
    length(unmatched_k),
    paste(unmatched_k, collapse = "\n")
  ))
}

krisper_item_urls <- krisper_joined |>
  filter(!is.na(url_item)) |>
  distinct(url_item)

# Step 2: Scrape item details — one request per item; cache: true avoids re-running
safe_results_k <- krisper_item_urls$url_item |>
  map(safely(get_item_details))

failed_k <- krisper_item_urls$url_item[map_lgl(safe_results_k, \(x) {
  !is.null(x$error)
})]
if (length(failed_k) > 0) {
  message(sprintf(
    "%d item(s) failed to parse:\n%s",
    length(failed_k),
    paste(failed_k, collapse = "\n")
  ))
}

df_item_details <- safe_results_k |> map("result") |> compact() |> list_rbind()

# Step 3: Find items where Krisper appears as a speaker in any stage
spoke_items <- df_item_details |>
  filter(!map_lgl(speeches, is.null)) |>
  unnest(speeches) |>
  filter(str_detect(speaker, "Krisper")) |>
  distinct(item_url)

# Step 4: One row per item (topics are item-level, replicated across stage rows)
df_krisper_spoke <- df_item_details |>
  semi_join(spoke_items, by = "item_url") |>
  distinct(item_url, .keep_all = TRUE)

# Visualise topics
df_krisper_spoke |>
  unnest(topics) |>
  count(topics, sort = TRUE) |>
  slice_head(n = 15) |>
  mutate(topics = fct_reorder(topics, n)) |>
  ggplot(aes(x = n, y = topics)) +
  geom_col(fill = "#216e39") +
  labs(
    x = "Number of items",
    y = NULL,
    title = "What did Krisper actually speak about?",
    subtitle = "Topics of parliamentary items where she personally contributed"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title.position = "plot",
    panel.grid.major.y = element_blank()
  )
```

The same analysis applies to any MP. Here is Ernst Gödl (ÖVP), offering a useful contrast to Krisper's profile:

```{r}
#| label: goedl-setup
#| eval: true
#| code-summary: "Gödl plenary speeches"
get_pad_intern("Gödl")

df_plenary_goedl <- get_mps_details(
  pad_intern = 83409,
  detail_type = "plenary",
  institution = "NR"
)

df_plenary_goedl |>
  unnest(position_name) |>
  count(position_name, sort = TRUE)
```

```{r}
#| label: fig-goedl-calendar
#| fig-cap: "GitHub-style calendar heatmap of Ernst Gödl's plenary speeches in the National Council."
#| eval: true
#| code-summary: "Calendar heatmap of Gödl's speeches"
#| fig-height: 7
#| fig-width: 10
gh_pal <- c(
  "0" = "#ebedf0",
  "1" = "#9be9a8",
  "2–3" = "#40c463",
  "4–6" = "#30a14e",
  "7+" = "#216e39"
)

date_min_g <- lubridate::floor_date(min(df_plenary_goedl$date), "year")
date_max_g <- lubridate::ceiling_date(max(df_plenary_goedl$date), "year") - 1

tibble(date = seq.Date(date_min_g, date_max_g, by = "day")) |>
  mutate(
    iso_year = lubridate::isoyear(date),
    week_no = lubridate::isoweek(date),
    dow = lubridate::wday(date, label = TRUE, abbr = TRUE, week_start = 1)
  ) |>
  left_join(count(df_plenary_goedl, date, name = "n"), by = "date") |>
  replace_na(list(n = 0)) |>
  mutate(
    n_cat = cut(
      n,
      breaks = c(-Inf, 0, 1, 3, 6, Inf),
      labels = names(gh_pal),
      right = TRUE
    )
  ) |>
  ggplot(aes(x = week_no, y = fct_rev(dow), fill = n_cat)) +
  geom_tile(width = 0.85, height = 0.85, colour = NA) +
  facet_wrap(~iso_year, ncol = 1, strip.position = "left") +
  scale_fill_manual(values = gh_pal, name = "Speeches", drop = FALSE) +
  scale_x_continuous(
    breaks = c(1, 5, 9, 14, 18, 22, 27, 31, 35, 40, 44, 48),
    labels = month.abb,
    expand = c(0.01, 0)
  ) +
  scale_y_discrete(breaks = c("Mon", "Wed", "Fri"), expand = c(0, 0.5)) +
  coord_fixed() +
  labs(
    x = NULL,
    y = NULL,
    title = "Ernst Gödl — plenary speeches",
    subtitle = "National Council"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title.position = "plot",
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    strip.text.y.left = element_text(angle = 0, size = 9, hjust = 1),
    legend.position = "bottom",
    legend.key.size = unit(0.55, "cm"),
    legend.direction = "horizontal",
    panel.spacing.y = unit(0.6, "cm"),
    plot.background = element_rect(fill = "white", colour = NA)
  )
```

```{r}
#| label: fig-goedl-topics
#| fig-cap: "Topics of parliamentary items where Ernst Gödl personally gave a speech."
#| eval: true
#| cache: true
#| code-summary: "Topics of items Gödl actually spoke on"
goedl_periods <- df_plenary_goedl |>
  distinct(legis_period) |>
  mutate(p = as.integer(as.roman(legis_period))) |>
  pull(p)

df_meeting_acts_g <- map(
  goedl_periods,
  \(p) {
    get_plenary_meetings(
      institution = "NR",
      legis_period = p,
      meeting_and_activities = "activities"
    )
  }
) |>
  list_rbind()

goedl_joined <- df_plenary_goedl |>
  distinct(meeting_url) |>
  mutate(
    url_meeting_norm = meeting_url |>
      str_replace("/sitzung/", "/gegenstand/") |>
      str_remove("\\?.*")
  ) |>
  left_join(df_meeting_acts_g, by = c("url_meeting_norm" = "url_meeting"))

unmatched_g <- goedl_joined |> filter(is.na(url_item)) |> pull(meeting_url)
if (length(unmatched_g) > 0) {
  message(sprintf(
    "%d meeting(s) unmatched in activities data:\n%s",
    length(unmatched_g),
    paste(unmatched_g, collapse = "\n")
  ))
}

goedl_item_urls <- goedl_joined |>
  filter(!is.na(url_item)) |>
  distinct(url_item)

safe_results_g <- goedl_item_urls$url_item |>
  map(safely(get_item_details))

failed_g <- goedl_item_urls$url_item[map_lgl(safe_results_g, \(x) {
  !is.null(x$error)
})]
if (length(failed_g) > 0) {
  message(sprintf(
    "%d item(s) failed to parse:\n%s",
    length(failed_g),
    paste(failed_g, collapse = "\n")
  ))
}

df_item_details_g <- safe_results_g |>
  map("result") |>
  compact() |>
  list_rbind()

spoke_items_g <- df_item_details_g |>
  filter(!map_lgl(speeches, is.null)) |>
  unnest(speeches) |>
  filter(str_detect(speaker, "Gödl")) |>
  distinct(item_url)

df_goedl_spoke <- df_item_details_g |>
  semi_join(spoke_items_g, by = "item_url") |>
  distinct(item_url, .keep_all = TRUE)

df_goedl_spoke |>
  unnest(topics) |>
  count(topics, sort = TRUE) |>
  slice_head(n = 15) |>
  mutate(topics = fct_reorder(topics, n)) |>
  ggplot(aes(x = n, y = topics)) +
  geom_col(fill = "#62C2CE") +
  labs(
    x = "Number of items",
    y = NULL,
    title = "What did Gödl actually speak about?",
    subtitle = "Topics of parliamentary items where he personally contributed"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title.position = "plot",
    panel.grid.major.y = element_blank()
  )
```

And finally August Wöginger (ÖVP), the long-serving parliamentary club chairman — his profile as a senior government-side MP should look quite different again:

```{r}
#| label: woeginger-setup
#| eval: true
#| code-summary: "Wöginger plenary speeches"
get_pad_intern("Wöginger")

df_plenary_woeginger <- get_mps_details(
  pad_intern = 14795,
  detail_type = "plenary",
  institution = "NR"
)

df_plenary_woeginger |>
  unnest(position_name) |>
  count(position_name, sort = TRUE)
```

```{r}
#| label: fig-woeginger-calendar
#| fig-cap: "GitHub-style calendar heatmap of August Wöginger's plenary speeches in the National Council."
#| eval: true
#| code-summary: "Calendar heatmap of Wöginger's speeches"
#| fig-width: 10
#| fig-height: 28
gh_pal <- c(
  "0" = "#ebedf0",
  "1" = "#9be9a8",
  "2–3" = "#40c463",
  "4–6" = "#30a14e",
  "7+" = "#216e39"
)

date_min_w <- lubridate::floor_date(min(df_plenary_woeginger$date), "year")
date_max_w <- lubridate::ceiling_date(max(df_plenary_woeginger$date), "year") -
  1

tibble(date = seq.Date(date_min_w, date_max_w, by = "day")) |>
  mutate(
    iso_year = lubridate::isoyear(date),
    week_no = lubridate::isoweek(date),
    dow = lubridate::wday(date, label = TRUE, abbr = TRUE, week_start = 1)
  ) |>
  left_join(count(df_plenary_woeginger, date, name = "n"), by = "date") |>
  replace_na(list(n = 0)) |>
  mutate(
    n_cat = cut(
      n,
      breaks = c(-Inf, 0, 1, 3, 6, Inf),
      labels = names(gh_pal),
      right = TRUE
    )
  ) |>
  ggplot(aes(x = week_no, y = fct_rev(dow), fill = n_cat)) +
  geom_tile(width = 0.85, height = 0.85, colour = NA) +
  facet_wrap(~iso_year, ncol = 1, strip.position = "left") +
  scale_fill_manual(values = gh_pal, name = "Speeches", drop = FALSE) +
  scale_x_continuous(
    breaks = c(1, 5, 9, 14, 18, 22, 27, 31, 35, 40, 44, 48),
    labels = month.abb,
    expand = c(0.01, 0)
  ) +
  scale_y_discrete(breaks = c("Mon", "Wed", "Fri"), expand = c(0, 0.5)) +
  labs(
    x = NULL,
    y = NULL,
    title = "August Wöginger — plenary speeches",
    subtitle = "National Council"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title.position = "plot",
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    strip.text.y.left = element_text(angle = 0, size = 9, hjust = 1),
    legend.position = "bottom",
    legend.key.size = unit(0.55, "cm"),
    legend.direction = "horizontal",
    panel.spacing.y = unit(0.6, "cm"),
    plot.background = element_rect(fill = "white", colour = NA)
  )
```

```{r}
#| label: fig-woeginger-topics
#| fig-cap: "Topics of parliamentary items where August Wöginger personally gave a speech."
#| eval: true
#| cache: true
#| code-summary: "Topics of items Wöginger actually spoke on"
woeginger_periods <- df_plenary_woeginger |>
  distinct(legis_period) |>
  mutate(p = as.integer(as.roman(legis_period))) |>
  pull(p)

df_meeting_acts_w <- map(
  woeginger_periods,
  \(p) {
    get_plenary_meetings(
      institution = "NR",
      legis_period = p,
      meeting_and_activities = "activities"
    )
  }
) |>
  list_rbind()

woeginger_joined <- df_plenary_woeginger |>
  distinct(meeting_url) |>
  mutate(
    url_meeting_norm = meeting_url |>
      str_replace("/sitzung/", "/gegenstand/") |>
      str_remove("\\?.*")
  ) |>
  left_join(df_meeting_acts_w, by = c("url_meeting_norm" = "url_meeting"))

unmatched_w <- woeginger_joined |> filter(is.na(url_item)) |> pull(meeting_url)
if (length(unmatched_w) > 0) {
  message(sprintf(
    "%d meeting(s) unmatched in activities data:\n%s",
    length(unmatched_w),
    paste(unmatched_w, collapse = "\n")
  ))
}

woeginger_item_urls <- woeginger_joined |>
  filter(!is.na(url_item)) |>
  distinct(url_item)

safe_results_w <- woeginger_item_urls$url_item |>
  map(safely(get_item_details))

failed_w <- woeginger_item_urls$url_item[map_lgl(safe_results_w, \(x) {
  !is.null(x$error)
})]
if (length(failed_w) > 0) {
  message(sprintf(
    "%d item(s) failed to parse:\n%s",
    length(failed_w),
    paste(failed_w, collapse = "\n")
  ))
}

df_item_details_w <- safe_results_w |>
  map("result") |>
  compact() |>
  list_rbind()

spoke_items_w <- df_item_details_w |>
  filter(!map_lgl(speeches, is.null)) |>
  unnest(speeches) |>
  filter(str_detect(speaker, "Wöginger")) |>
  distinct(item_url)

df_woeginger_spoke <- df_item_details_w |>
  semi_join(spoke_items_w, by = "item_url") |>
  distinct(item_url, .keep_all = TRUE)

df_woeginger_spoke |>
  unnest(topics) |>
  count(topics, sort = TRUE) |>
  slice_head(n = 15) |>
  mutate(topics = fct_reorder(topics, n)) |>
  ggplot(aes(x = n, y = topics)) +
  geom_col(fill = "#62C2CE") +
  labs(
    x = "Number of items",
    y = NULL,
    title = "What did Wöginger actually speak about?",
    subtitle = "Topics of parliamentary items where he personally contributed"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title.position = "plot",
    panel.grid.major.y = element_blank()
  )
```

With data on all three MPs in hand, we can compare their topical focus side by side. The chart below shows the relative composition of topics for items each MP personally spoke on, with the top 4 topics per MP labelled and everything else grouped as "Other".

```{r}
#| label: fig-mp-topics-comparison
#| fig-cap: "Relative topical composition of parliamentary items each MP personally spoke on. Top 4 topics per MP shown individually; all others grouped as 'Other'."
#| eval: true
#| code-summary: "Comparative topic composition across MPs"
#| fig-height: 4
#| fig-width: 10
# Combine topic counts for all three MPs
df_all_topics <- bind_rows(
  df_krisper_spoke |> unnest(topics) |> count(topics) |> mutate(mp = "Krisper"),
  df_goedl_spoke |> unnest(topics) |> count(topics) |> mutate(mp = "Gödl"),
  df_woeginger_spoke |>
    unnest(topics) |>
    count(topics) |>
    mutate(mp = "Wöginger")
)

# Per-MP top 4 topics
top4_per_mp <- df_all_topics |>
  group_by(mp) |>
  slice_max(n, n = 4, with_ties = FALSE) |>
  ungroup() |>
  distinct(mp, topics)

# Relabel non-top-4 as "Other" and aggregate
df_plot <- df_all_topics |>
  left_join(top4_per_mp |> mutate(keep = TRUE), by = c("mp", "topics")) |>
  mutate(topic_label = if_else(is.na(keep), "Other", topics)) |>
  summarise(n = sum(n), .by = c(mp, topic_label)) |>
  mutate(pct = n / sum(n), .by = mp)

# Order named topics by combined frequency; "Other" always last
topic_order <- df_plot |>
  filter(topic_label != "Other") |>
  summarise(total = sum(n), .by = topic_label) |>
  arrange(desc(total)) |>
  pull(topic_label)

df_plot <- df_plot |>
  mutate(
    topic_label = factor(topic_label, levels = c(topic_order, "Other")),
    mp = factor(mp, levels = c("Wöginger", "Gödl", "Krisper"))
  )

# Colour palette — interpolated Set2 for named topics, grey for Other
n_named <- length(topic_order)
topic_colors <- setNames(
  c(colorRampPalette(RColorBrewer::brewer.pal(8, "Set2"))(n_named), "#cccccc"),
  c(topic_order, "Other")
)

ggplot(df_plot, aes(x = mp, y = pct, fill = topic_label)) +
  geom_col(position = position_stack(reverse = TRUE), width = 0.6) +
  geom_text(
    aes(label = if_else(pct >= 0.05, scales::percent(pct, accuracy = 1), "")),
    position = position_stack(vjust = 0.5, reverse = TRUE),
    size = 3,
    colour = "white",
    fontface = "bold"
  ) +
  scale_fill_manual(values = topic_colors, name = NULL) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  coord_flip() +
  labs(
    x = NULL,
    y = NULL,
    title = "Topical focus of plenary speeches",
    subtitle = "Top 4 topics per MP; remaining topics grouped as 'Other'"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title.position = "plot",
    legend.position = "bottom",
    legend.key.size = unit(0.45, "cm"),
    panel.grid.major.y = element_blank()
  ) +
  guides(fill = guide_legend(nrow = 2))
```

```{r}
#| label: all-items-fetch
#| cache: true
df_all_items <- get_items(legis_period = seq(20, 27))
```

```{r}
#| label: all-item-details-fetch
#| cache: true
item_urls_all <- df_all_items |> distinct(item_url) |> pull(item_url)

safe_results_all <- item_urls_all |> map(safely(get_item_details))

failed_all <- item_urls_all[map_lgl(safe_results_all, \(x) !is.null(x$error))]
if (length(failed_all) > 0) {
  message(sprintf(
    "%d item(s) failed to parse:\n%s",
    length(failed_all),
    paste(failed_all, collapse = "\n")
  ))
}

df_all_item_details <- safe_results_all |>
  map("result") |>
  compact() |>
  list_rbind()
```

```{r}
#| label: export-rds
#| include: false
saveRDS(df_all_items, "df_all_items.rds")
saveRDS(df_all_item_details, "df_all_item_details.rds")
```

```{r}
#| label: fig-topics-by-year
#| fig-cap: "Topical composition of all National Council items by year"
#| fig-height: 6
df_yearly <- df_all_items |>
  mutate(year = lubridate::year(date)) |>
  unnest(topics) |>
  count(year, topics)

top_topics <- df_yearly |>
  summarise(total = sum(n), .by = topics) |>
  slice_max(total, n = 6, with_ties = FALSE) |>
  pull(topics)

df_plot_yearly <- df_yearly |>
  mutate(topic_label = if_else(topics %in% top_topics, topics, "Other")) |>
  summarise(n = sum(n), .by = c(year, topic_label))

topic_order <- df_plot_yearly |>
  filter(topic_label != "Other") |>
  summarise(total = sum(n), .by = topic_label) |>
  arrange(desc(total)) |>
  pull(topic_label)

df_plot_yearly <- df_plot_yearly |>
  mutate(topic_label = factor(topic_label, levels = c(topic_order, "Other")))

n_named <- length(topic_order)
topic_colors <- setNames(
  c(colorRampPalette(RColorBrewer::brewer.pal(8, "Set2"))(n_named), "#cccccc"),
  c(topic_order, "Other")
)

ggplot(df_plot_yearly, aes(x = year, y = n, fill = topic_label)) +
  geom_col(position = position_stack(reverse = TRUE), width = 0.75) +
  scale_fill_manual(values = topic_colors, name = NULL) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = NULL,
    y = "Number of items",
    title = "How the topical focus of parliamentary work changed over time",
    subtitle = "All item types, National Council, periods XX–XXVII; top 6 topics shown"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title.position = "plot",
    legend.position = "bottom",
    legend.key.size = unit(0.45, "cm"),
    panel.grid.major.x = element_blank()
  ) +
  guides(fill = guide_legend(nrow = 2))
```

```{r}
#| label: fig-topics-by-year-weighted
#| fig-cap: "Topical composition of National Council speeches by year (speech-weighted)"
#| fig-height: 6
df_yearly_weighted <- df_all_item_details |>
  mutate(
    year = lubridate::year(date_introduced),
    n_speeches = map_int(speeches, \(s) if (is.null(s)) 0L else nrow(s))
  ) |>
  filter(n_speeches > 0) |>
  unnest(topics) |>
  group_by(year, topics) |>
  summarise(n = sum(n_speeches), .groups = "drop")

df_plot_yearly_weighted <- df_yearly_weighted |>
  mutate(topic_label = if_else(topics %in% top_topics, topics, "Other")) |>
  summarise(n = sum(n), .by = c(year, topic_label)) |>
  mutate(topic_label = factor(topic_label, levels = c(topic_order, "Other")))

ggplot(df_plot_yearly_weighted, aes(x = year, y = n, fill = topic_label)) +
  geom_col(position = position_stack(reverse = TRUE), width = 0.75) +
  scale_fill_manual(values = topic_colors, name = NULL) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = NULL,
    y = "Number of speeches",
    title = "How the intensity of parliamentary debate changed over time",
    subtitle = "Items weighted by speeches held; all item types, National Council, periods XX–XXVII"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title.position = "plot",
    legend.position = "bottom",
    legend.key.size = unit(0.45, "cm"),
    panel.grid.major.x = element_blank()
  ) +
  guides(fill = guide_legend(nrow = 2))
```

## Speech-level analysis

The charts above treat parliamentary items as the unit of analysis. But `get_plenary_meeting_details()` lets us go one level deeper — to individual speeches — and links each speech to the speaker, duration, and debate context. Combining this with `get_mps()` for gender and party affiliation produces a speech-level dataset that opens up a different class of questions: not just *what* was debated, but *who spoke*, *how long*, and *on what topic*.

The dataset covers all 90,610 speeches delivered in the National Council across legislative periods XX–XXVII (1996–2024). Each row records the speaker's name, gender, party, the date and duration of the speech (in seconds), the debate type, and — where a formal resolution was adopted — the item-level topic classifications (topics, EuroVoc descriptors, keywords).

```{r}
#| label: speech-data-load
#| include: false
df_speeches <- readRDS("df_speech_dataset.rds")
```

### Do women and men speak for different lengths of time by topic?

A first cut: average speech duration by gender and topic, limited to speeches with matched topic data.

```{r}
#| label: fig-speech-duration-gender
#| fig-cap: "Mean speech duration by gender and topic"
#| fig-height: 6
df_speeches |>
  filter(!is.na(gender), !map_lgl(topics, is.null), duration_sec > 10) |>
  unnest(topics) |>
  group_by(topics, gender) |>
  summarise(mean_sec = mean(duration_sec), n = n(), .groups = "drop") |>
  filter(n >= 30) |>
  pivot_wider(names_from = gender, values_from = c(mean_sec, n)) |>
  filter(!is.na(mean_sec_female), !is.na(mean_sec_male)) |>
  mutate(
    diff = mean_sec_female - mean_sec_male,
    topics = fct_reorder(topics, diff)
  ) |>
  ggplot(aes(x = diff, y = topics)) +
  geom_col(aes(fill = diff > 0), width = 0.6, show.legend = FALSE) +
  geom_vline(xintercept = 0, linewidth = 0.4) +
  scale_fill_manual(values = c("TRUE" = "#4DAF4A", "FALSE" = "#E41A1C")) +
  scale_x_continuous(labels = \(x) {
    paste0(ifelse(x > 0, "+", ""), round(x), "s")
  }) +
  labs(
    x = "Women − Men (seconds)",
    y = NULL,
    title = "Women speak longer on gender equality and family topics",
    subtitle = "Raw difference in mean duration; positive = women speak longer"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title.position = "plot", panel.grid.major.y = element_blank())
```

The raw picture shows women speaking longer on *Frauen und Gleichbehandlung* (+29 s), *Innovation* (+23 s), and *Familie* (+10 s), while men speak longer on *Landesverteidigung* (−18 s), *Europäische Union* (−18 s), and *Budget* (−13 s). But raw means conflate individual speaking style with topic effects — a handful of prolific male MPs dominating certain debates could drive the whole pattern.

### Controlling for individual speaking style: a mixed effects model

To isolate the topic-specific gender effect we fit a log-linear mixed effects model with a random intercept per MP (absorbing individual baseline speaking style), fixed effects for party, legislative period, debate type, and formal time limit:

$$\log(\text{duration}) = \beta_0 + \beta_{\text{gender}} \times \text{topic} + \beta_{\text{party}} + \beta_{\text{period}} + \beta_{\text{type}} + \log(\text{limit}) + u_{\text{MP}} + \varepsilon$$

```{r}
#| label: speech-model-gender
#| cache: true
#| message: false
library(nlme)

df_model_gender <- df_speeches |>
  filter(!is.na(gender), !map_lgl(topics, is.null), duration_sec > 10) |>
  mutate(
    limit_sec = suppressWarnings(as.integer(speech_limit) * 60L),
    log_limit = if_else(
      !is.na(limit_sec) & limit_sec > 0,
      log(limit_sec),
      NA_real_
    ),
    log_dur = log(duration_sec),
    gender = factor(gender, levels = c("female", "male")),
    debate_type = factor(debate_type),
    pad_intern = factor(pad_intern)
  ) |>
  unnest(topics) |>
  mutate(topics = factor(topics)) |>
  filter(!is.na(log_limit))

fit_gender <- lme(
  log_dur ~ gender * topics + parl_group + period_int + debate_type + log_limit,
  random = ~ 1 | pad_intern,
  data = df_model_gender,
  method = "REML",
  na.action = na.omit,
  control = lmeControl(opt = "optim", maxIter = 200)
)

# Extract tTable here while nlme is loaded; cached alongside the model
tt_gender <- as.data.frame(summary(fit_gender)$tTable)
tt_gender$term <- rownames(tt_gender)
```

```{r}
#| label: fig-speech-model-gender
#| fig-cap: "Gender × topic interactions from mixed effects model"
#| fig-height: 6
# tt_gender is pre-extracted in the cached speech-model-gender chunk

# Overall gender effect (on reference topic)
main_gender <- tt_gender["gendermale", ]
cat(sprintf(
  "Overall gender effect: β = %.3f, SE = %.3f, t = %.2f, p = %.3f\n",
  main_gender$Value,
  main_gender$`Std.Error`,
  main_gender$`t-value`,
  main_gender$`p-value`
))

# Gender:topic interactions
gi <- tt_gender[grep("gendermale:topics", tt_gender$term), ]
gi$topic <- sub("gendermale:topics", "", gi$term)
gi$total_effect <- main_gender$Value + gi$Value # total gender gap on this topic
gi$topic <- fct_reorder(gi$topic, gi$total_effect)

ggplot(gi, aes(x = total_effect, y = topic)) +
  geom_vline(xintercept = 0, linewidth = 0.4) +
  geom_vline(
    xintercept = main_gender$Value,
    linetype = "dashed",
    colour = "grey50",
    linewidth = 0.4
  ) +
  geom_pointrange(
    aes(
      xmin = total_effect - 1.96 * `Std.Error`,
      xmax = total_effect + 1.96 * `Std.Error`,
      colour = `p-value` < 0.05
    ),
    size = 0.4
  ) +
  scale_colour_manual(
    values = c("TRUE" = "#E41A1C", "FALSE" = "grey40"),
    name = "p < 0.05"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Total gender effect on log(duration) — male vs female",
    y = NULL,
    title = "After controlling for MP style, topic-specific gender gaps vanish",
    subtitle = "Dashed line = overall gender effect; red = significant at 5%"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title.position = "plot", panel.grid.major.y = element_blank())
```

After accounting for individual speaking style, the apparent topic-level gender differences largely disappear. The overall effect is small but significant (β = −0.037, p = 0.046): on average women speak ~3.7% longer than men after controls. Only *Budget und Finanzen* shows a significant interaction (p = 0.037), where the gender gap narrows almost to zero — men speak nearly as long as women on budget topics, unlike most other areas where women are slightly longer.

### Does party predict how long MPs speak by topic?

Replacing gender with party affiliation tests whether ideological positioning shapes topic-specific speaking time. We restrict to the six parties with sufficient observations (ÖVP, SPÖ, FPÖ, GRÜNE, NEOS, BZÖ) and use ÖVP as the reference category.

```{r}
#| label: speech-model-party
#| cache: true
#| message: false
harmonise_party <- function(x) {
  case_when(
    str_detect(x, "Volkspartei") ~ "ÖVP",
    str_detect(x, "Sozialdemokratis") ~ "SPÖ",
    str_detect(x, "Freiheitlich") & str_detect(x, "BZÖ") ~ "FPÖ/BZÖ",
    str_detect(x, "Freiheitlich") ~ "FPÖ",
    str_detect(x, "Grün") ~ "GRÜNE",
    str_detect(x, "NEOS") ~ "NEOS",
    str_detect(x, "BZÖ") ~ "BZÖ",
    TRUE ~ NA_character_
  )
}

df_model_party <- df_speeches |>
  filter(!is.na(parl_group), !map_lgl(topics, is.null), duration_sec > 10) |>
  mutate(
    party = harmonise_party(parl_group),
    limit_sec = suppressWarnings(as.integer(speech_limit) * 60L),
    log_limit = if_else(
      !is.na(limit_sec) & limit_sec > 0,
      log(limit_sec),
      NA_real_
    ),
    log_dur = log(duration_sec)
  ) |>
  filter(!is.na(party), !is.na(log_limit)) |>
  unnest(topics) |>
  mutate(
    party = factor(
      party,
      levels = c("ÖVP", "SPÖ", "FPÖ", "GRÜNE", "NEOS", "BZÖ")
    ),
    topics = factor(topics),
    debate_type = factor(debate_type),
    pad_intern = factor(pad_intern)
  )

fit_party <- lme(
  log_dur ~ party * topics + gender + period_int + debate_type + log_limit,
  random = ~ 1 | pad_intern,
  data = df_model_party,
  method = "REML",
  na.action = na.omit,
  control = lmeControl(opt = "optim", maxIter = 300)
)

# Extract tTable here while nlme is loaded; cached alongside the model
tt_party <- as.data.frame(summary(fit_party)$tTable)
tt_party$term <- rownames(tt_party)
```

```{r}
#| label: fig-speech-model-party
#| fig-cap: "Main party effects on speech duration (vs ÖVP)"
#| fig-height: 4
# tt_party is pre-extracted in the cached speech-model-party chunk

main_party <- tt_party[
  grepl("^party[A-Z]", tt_party$term) & !grepl(":", tt_party$term),
]
main_party$party <- str_remove(main_party$term, "party")
main_party$party <- fct_reorder(main_party$party, main_party$Value)

ggplot(main_party, aes(x = Value, y = party)) +
  geom_vline(xintercept = 0, linewidth = 0.4) +
  geom_pointrange(
    aes(
      xmin = Value - 1.96 * `Std.Error`,
      xmax = Value + 1.96 * `Std.Error`,
      colour = `p-value` < 0.05
    ),
    size = 0.5
  ) +
  scale_colour_manual(
    values = c("TRUE" = "#E41A1C", "FALSE" = "grey40"),
    name = "p < 0.05"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Effect on log(duration) vs ÖVP",
    y = NULL,
    title = "SPÖ speaks significantly shorter than ÖVP across all topics",
    subtitle = "Mixed effects model controlling for MP style, period, debate type, time limit"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title.position = "plot", panel.grid.major.y = element_blank())
```

SPÖ is the only party that speaks significantly shorter than ÖVP (β = −0.056, p = 0.014, ~5.4% on the log scale). No other party differs significantly from ÖVP, and — crucially — none of the 110 party × topic interaction terms reach significance. The formal *Redezeit* allocation system, which distributes floor time proportionally to party size, appears to dominate over any ideological differences in how parties allocate their speaking time across topics.

### Do female MPs speak at their representation rate?

The gender gap in speech duration is one question; a different question is whether female MPs speak at all at the rate their numbers would predict. A party where 30% of MPs are women should, all else equal, see roughly 30% of its plenary speeches delivered by women. If the observed share is lower, female MPs are speaking below their representation rate — a signal of internal party culture, not just headcount.

The analysis pairs the MP roster from `get_mps()` with the speech dataset to compute, for each party × legislative period, the ratio of speeches-per-female-MP to speeches-per-male-MP. A ratio of 1 means parity; below 1 means female MPs speak less per capita than their male colleagues.

```{r}
#| label: female-rate-data
#| cache: true
PARTIES_6 <- c("ÖVP", "SPÖ", "FPÖ", "GRÜNE", "NEOS", "BZÖ")

df_mps_raw <- readRDS("_cache_mps_all.rds")

df_roster <- df_mps_raw |>
  unnest(mp_details) |>
  ungroup() |>
  mutate(
    period_int = suppressWarnings(as.integer(legis_period)) |>
      (\(v) if_else(is.na(v), as.integer(as.roman(legis_period)), v))(),
    party = harmonise_party(parl_group)
  ) |>
  filter(!is.na(party), !is.na(gender), party %in% PARTIES_6) |>
  distinct(pad_intern, period_int, party, gender) |>
  ungroup() |>
  count(party, period_int, gender, name = "n_mps")

df_spk_counts <- df_speeches |>
  filter(!is.na(gender), !is.na(parl_group), duration_sec > 10) |>
  mutate(party = harmonise_party(parl_group)) |>
  filter(!is.na(party), party %in% PARTIES_6) |>
  count(party, period_int, gender, name = "n_speeches")

df_ratio <- df_roster |>
  left_join(df_spk_counts, by = c("party", "period_int", "gender")) |>
  replace_na(list(n_speeches = 0L)) |>
  mutate(rate = n_speeches / n_mps) |>
  pivot_wider(names_from = gender, values_from = c(n_mps, rate, n_speeches)) |>
  filter(
    !is.na(rate_female),
    !is.na(rate_male),
    rate_male > 0,
    n_mps_female >= 3
  ) |>
  mutate(
    activity_ratio = rate_female / rate_male,
    female_mp_share = n_mps_female / (n_mps_female + n_mps_male),
    female_speech_share = n_speeches_female /
      (n_speeches_female + n_speeches_male),
    period_label = paste0("GP ", as.roman(period_int))
  )
```

```{r}
#| label: fig-female-rate-scatter
#| fig-cap: "Female MP share vs female speech share by party and legislative period"
#| fig-height: 5.5
party_colors_6 <- c(
  ÖVP = "#62a8e5",
  SPÖ = "#E41A1C",
  FPÖ = "#2171b5",
  GRÜNE = "#4daf4a",
  NEOS = "#fa9fb5",
  BZÖ = "#ff7f00"
)

ggplot(
  df_ratio,
  aes(
    x = female_mp_share,
    y = female_speech_share,
    colour = party,
    label = period_label
  )
) +
  geom_abline(
    slope = 1,
    intercept = 0,
    linetype = "dashed",
    colour = "grey50",
    linewidth = 0.5
  ) +
  geom_point(size = 2.5, alpha = 0.85) +
  geom_text(size = 2.2, vjust = -0.7, show.legend = FALSE) +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA),
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA),
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  scale_colour_manual(values = party_colors_6, name = NULL) +
  labs(
    x = "Female share of MPs",
    y = "Female share of plenary speeches",
    title = "Are female MPs speaking at their representation rate?",
    subtitle = "Each point = one party × legislative period; diagonal = parity (expected = actual)"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title.position = "plot", legend.position = "bottom")
```

```{r}
#| label: fig-female-rate-ratio
#| fig-cap: "Female-to-male speech activity ratio by party across legislative periods"
#| fig-height: 5
df_ratio |>
  mutate(party = factor(party, levels = PARTIES_6)) |>
  ggplot(aes(x = period_int, y = activity_ratio, colour = party)) +
  geom_hline(
    yintercept = 1,
    linetype = "dashed",
    colour = "grey50",
    linewidth = 0.5
  ) +
  geom_line(linewidth = 0.8, alpha = 0.8) +
  geom_point(size = 2.5) +
  scale_x_continuous(
    breaks = 20:27,
    labels = paste0("GP ", as.roman(20:27))
  ) +
  scale_y_continuous(labels = scales::label_number(suffix = "×")) +
  scale_colour_manual(values = party_colors_6, name = NULL) +
  labs(
    x = NULL,
    y = "Speeches per female MP / speeches per male MP",
    title = "FPÖ women speak the least relative to their numbers; ÖVP women the most",
    subtitle = "Ratio = 1 means parity; above 1 means female MPs speak more per capita than male MPs"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title.position = "plot",
    axis.text.x = element_text(size = 8),
    legend.position = "bottom"
  )
```

The pattern is consistent across periods. GRÜNE women track close to or above parity in every period — the party's internal culture appears to actively support female participation. ÖVP women also sit close to the diagonal, though slightly below. SPÖ shows a modest but persistent gap that has widened in recent periods.

FPÖ is the outlier in the other direction. In every period with sufficient female MPs, FPÖ women speak at well below half the rate of their male colleagues. Period XXII is extreme: the post-BZÖ split left a small rump group, and the female MPs in it appear to have delivered no speeches at all. NEOS entered parliament in period XXV with a marked gap that has since narrowed — plausible for a new party still establishing internal speaking norms.

The parity line on the scatter makes the point directly: nearly all SPÖ, ÖVP, and GRÜNE points cluster around the diagonal; FPÖ points sit well below it regardless of how many women the party has in parliament.

### Parliamentary activities

```{r}
#| eval: true
#| code-summary: "Parliamentary activities"
# Written questions submitted by an MP

get_pad_intern("Pilz")

df_activities <- get_mps_details(
  pad_intern = 1210,
  detail_type = "activities",
  institution = "NR",
  item = "JMIN" # Written questions to the Federal Government
)

nrow(df_activities)

glimpse(df_activities)
```

## Committees (Ausschüsse)

`get_committees()` retrieves committee information for a given legislative period and institution. It can also extract membership lists when `details_type = "members"` is set.

### Listing committees

```{r}
#| eval: true
#| code-summary: "List committees"
# All committees of the National Council in the 27th period
committees_nr27 <- get_committees(legis_period = 27, institution = "NR")
committees_nr27
```

### Committees of inquiry (Untersuchungsausschüsse)

The `citation` parameter allows filtering by committee type code. Committees of inquiry use the code "USA":

```{r}
#| eval: true
#| code-summary: "Committees of inquiry"
# How many committees of inquiry per legislative period?
inquiry_committees <- map(
  seq(20, 28),
  \(x) get_committees(legis_period = x, institution = "NR", citation = "USA")
) |>
  list_rbind()

inquiry_committees |>
  mutate(
    legis_period = factor(legis_period, levels = as.character(as.roman(20:28)))
  ) |>
  count(legis_period, .drop = FALSE)
```

## Plenary meetings

`get_plenary_meetings()` retrieves information about plenary meetings from the Austrian Parliament's API. Data is available from the 20th legislative period onward, and covers the National Council (NR), the Federal Council (BR), and the Federal Assembly (BV).

The `meeting_and_activities` parameter controls what is returned:

- `"meetings"`: Meeting metadata — dates, meeting numbers, agenda URLs, and meeting type
- `"activities"`: Items that occurred during meetings — motions, questions, declarations, etc.

Additional filters include `session_type` (`"N"` for ordinary, `"A"` for extraordinary sessions) and `meeting_type` (`"S"`, `"SO"`, `"ZU"`, `"N"`).

```{r}
#| eval: true
#| code-summary: "Plenary meetings"
# National Council meetings in the 28th legislative period
meetings_28 <- get_plenary_meetings(
  institution = "NR",
  legis_period = 28,
  meeting_and_activities = "meetings"
)

glimpse(meetings_28)

# Count meetings per legislative period across multiple periods
meetings_multi <- get_plenary_meetings(
  institution = "NR",
  legis_period = c(26, 27),
  meeting_and_activities = "meetings"
)

meetings_multi |> count(legis_period)
```

```{r}
#| eval: true
#| code-summary: "Activities during plenary meetings"
# Parliamentary activities during 27th period meetings
activities_27 <- get_plenary_meetings(
  institution = "NR",
  legis_period = 27,
  meeting_and_activities = "activities"
)

glimpse(activities_27)
```

The Federal Assembly (Bundesversammlung), which convenes for presidential inaugurations and similar joint meetings, can be queried by setting `institution = "BV"` and `legis_period = NULL`.

## Transcripts (Stenographische Protokolle)

`get_transcripts()` provides access to stenographic protocols. Results include links to both HTML and PDF versions of the transcripts. For bulk downloads, the function supports PDF export

```{r}
#| eval: false
#| code-summary: "Search transcripts"
# Transcripts mentioning "Gesundheit" in the 28th period
transcripts <- get_transcripts(
  search_string = "gesundheit",
  legis_period = 28,
  session_type = "NRSITZ",
  echo = TRUE
)

glimpse(transcripts)
```

When defining a destination folder, the function provides for the download of the transcript files.

```{r}
#| eval: false
#| code-summary: "Download transcript PDFs"
# Download PDFs to a local folder
get_transcripts(
  legis_period = 28,
  session_type = "NRSITZ",
  export = "pdf",
  export_destination = "my_transcripts"
)
```

## The `echo` argument

Most functions accept an `echo` argument. When set to `TRUE`, the function prints the search parameters and a direct URL linking to the same results on the Parliament's website:

```{r}
#| eval: false
#| code-summary: "Echo parameter example"
get_items(item = "RV", legis_period = 27, echo = TRUE)
#> {"NRBR":"NR","GP":"XXVII.","VHG":"RV"}
#> https://www.parlament.gv.at/recherchieren/gegenstaende/...
#> [1] 245
```

I find this useful for verification, for quickly sharing a specific query with someone who prefers the web interface, and for sanity-checking that the API query is doing what I expect. It's a small feature but it's saved me more than once.

## FPÖ: A party in transformation

Few parties in post-war Austrian history have undergone as dramatic a trajectory as the FPÖ. From a minor liberal-nationalist party in the 1950s, it was reshaped into a populist right-wing force under Jörg Haider in the 1980s and 1990s, entered government twice (2000–2005 and 2017–2019), survived an internal split that spawned the BZÖ, and re-emerged as the dominant opposition force after the 2019 Ibiza affair ended its second coalition. The ParlAT data — spanning legislative periods XX–XXVII (2002–2024) — captures the arc from post-Haider peak through opposition years, the Türkis-Blau government, and back into opposition.

The party group name itself changed several times: "Klub der Freiheitlichen Partei Österreichs" in periods XX–XXI, "Freiheitlicher Parlamentsklub - BZÖ" in the first half of period XXII (before the April 2005 split), then "Freiheitlicher Parlamentsklub" from the split onwards through period XXVII. The analysis below treats all of these as "FPÖ", excluding the BZÖ breakaway faction which formed its own parliamentary group.

```{r}
#| label: fpoe-data-prep
#| cache: true
df_fpoe <- df_speeches |>
  filter(!is.na(parl_group), duration_sec > 10) |>
  mutate(
    is_fpoe = str_detect(parl_group, "Freiheitlich") &
      !str_detect(parl_group, "BZÖ"),
    year = as.integer(format(date, "%Y"))
  )

# Legislative period labels (approx. start years)
period_labels <- tibble(
  period_int = 20:27,
  period_label = paste0("GP ", as.roman(20:27)),
  year_start = c(2002, 2003, 2006, 2008, 2013, 2017, 2019, 2024)
)

# Government annotations
govt_spans <- tibble(
  xmin = c(2002.0, 2017.5),
  xmax = c(2005.3, 2019.5),
  label = c("FPÖ in government\n(ÖVP–FPÖ I)", "FPÖ in government\n(ÖVP–FPÖ II)")
)
```

```{r}
#| label: fig-fpoe-presence
#| fig-cap: "FPÖ parliamentary speech volume by year, 2002–2024"
#| fig-height: 5
df_fpoe |>
  filter(is_fpoe, !is.na(year)) |>
  count(year) |>
  ggplot(aes(x = year, y = n)) +
  annotate(
    "rect",
    xmin = govt_spans$xmin,
    xmax = govt_spans$xmax,
    ymin = -Inf,
    ymax = Inf,
    fill = "#2171b5",
    alpha = 0.10
  ) +
  annotate(
    "text",
    x = (govt_spans$xmin + govt_spans$xmax) / 2,
    y = c(2700, 2700),
    label = govt_spans$label,
    size = 3,
    colour = "#2171b5",
    lineheight = 0.9
  ) +
  geom_vline(
    xintercept = 2019.35,
    linetype = "dashed",
    colour = "#cb181d",
    linewidth = 0.6
  ) +
  annotate(
    "text",
    x = 2019.55,
    y = 2600,
    label = "Ibiza\nscandal",
    hjust = 0,
    size = 3,
    colour = "#cb181d",
    lineheight = 0.9
  ) +
  geom_col(fill = "#2171b5", alpha = 0.7, width = 0.75) +
  scale_x_continuous(breaks = seq(2002, 2024, 2)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = NULL,
    y = "Speeches",
    title = "FPÖ speaks less when in government — opposition is its natural habitat",
    subtitle = "Annual number of FPÖ plenary speeches, National Council, 2002–2024"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title.position = "plot", panel.grid.major.x = element_blank())
```

When FPÖ is in government, its speech volume drops sharply — ministers speak for the government, and coalition discipline constrains backbencher activity. The pattern repeats across both coalition periods. The post-Ibiza years saw a rebound as the party returned to full-throated opposition.

```{r}
#| label: fpoe-topics-prep
#| cache: true
# Topics per legislative period, FPÖ only
df_fpoe_topics <- df_fpoe |>
  filter(is_fpoe, !map_lgl(topics, is.null), !is.na(period_int)) |>
  unnest(topics) |>
  count(period_int, topics)

top_topics_fpoe <- df_fpoe_topics |>
  summarise(total = sum(n), .by = topics) |>
  slice_max(total, n = 7, with_ties = FALSE) |>
  pull(topics)

df_fpoe_pct <- df_fpoe_topics |>
  mutate(topic_label = if_else(topics %in% top_topics_fpoe, topics, "Other")) |>
  summarise(n = sum(n), .by = c(period_int, topic_label)) |>
  group_by(period_int) |>
  mutate(pct = n / sum(n)) |>
  ungroup()

topic_order_fpoe <- df_fpoe_pct |>
  filter(topic_label != "Other") |>
  summarise(total = sum(n), .by = topic_label) |>
  arrange(desc(total)) |>
  pull(topic_label)

df_fpoe_pct <- df_fpoe_pct |>
  mutate(
    topic_label = factor(topic_label, levels = c(topic_order_fpoe, "Other"))
  )

n_named_f <- length(topic_order_fpoe)
fpoe_colors <- setNames(
  c(
    colorRampPalette(RColorBrewer::brewer.pal(8, "Set2"))(n_named_f),
    "#cccccc"
  ),
  c(topic_order_fpoe, "Other")
)

period_x_labels <- c(
  "20" = "GP XX\n2002–03",
  "21" = "GP XXI\n2003–06",
  "22" = "GP XXII\n2006–08",
  "23" = "GP XXIII\n2008–13",
  "24" = "GP XXIV\n2013–17",
  "25" = "GP XXV\n2017–19",
  "26" = "GP XXVI\n2019–24",
  "27" = "GP XXVII\n2024–"
)
```

```{r}
#| label: fig-fpoe-topic-composition
#| fig-cap: "FPÖ topic composition by legislative period"
#| fig-height: 6
ggplot(df_fpoe_pct, aes(x = factor(period_int), y = pct, fill = topic_label)) +
  geom_col(position = position_stack(reverse = TRUE), width = 0.8) +
  scale_x_discrete(labels = period_x_labels) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_manual(values = fpoe_colors, name = NULL) +
  labs(
    x = NULL,
    y = "Share of matched speeches",
    title = "How FPÖ's topical focus shifted across legislative periods",
    subtitle = "Share of FPÖ plenary speeches per topic group; top 7 topics shown"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title.position = "plot",
    legend.position = "bottom",
    legend.key.size = unit(0.45, "cm"),
    panel.grid.major.x = element_blank()
  ) +
  guides(fill = guide_legend(nrow = 2))
```

```{r}
#| label: fig-fpoe-overindex
#| fig-cap: "FPÖ topic over-representation relative to parliament average"
#| fig-height: 6
# For each topic: FPÖ share of all speeches vs. overall FPÖ share
fpoe_share_overall <- df_fpoe |>
  filter(!map_lgl(topics, is.null)) |>
  unnest(topics) |>
  summarise(
    n_fpoe = sum(is_fpoe, na.rm = TRUE),
    n_total = n(),
    .by = topics
  ) |>
  mutate(
    fpoe_share = n_fpoe / n_total,
    overall_fpoe = sum(n_fpoe) / sum(n_total),
    over_index = fpoe_share / overall_fpoe
  ) |>
  filter(n_total >= 100) |>
  arrange(desc(over_index))

top_n_show <- 18

df_oi <- bind_rows(
  slice_max(fpoe_share_overall, over_index, n = top_n_show %/% 2),
  slice_min(fpoe_share_overall, over_index, n = top_n_show %/% 2)
) |>
  distinct(topics, .keep_all = TRUE) |>
  mutate(
    dir = if_else(over_index >= 1, "over", "under"),
    label = topics
  )

ggplot(
  df_oi,
  aes(x = over_index - 1, y = reorder(label, over_index), fill = dir)
) +
  geom_col(width = 0.7, alpha = 0.85) +
  geom_vline(xintercept = 0, linewidth = 0.4) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_manual(
    values = c(over = "#cb181d", under = "#2171b5"),
    guide = "none"
  ) +
  labs(
    x = "Over-/under-representation vs. overall FPÖ speech share",
    y = NULL,
    title = "FPÖ dominates security, migration and EU topics; avoids social policy",
    subtitle = "Topics where FPÖ's share of all speeches deviates most from its overall share"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title.position = "plot", panel.grid.major.y = element_blank())
```

```{r}
#| label: fig-fpoe-govt-vs-oppo
#| fig-cap: "FPÖ topic profile: government vs. opposition periods"
#| fig-height: 6
# Government periods: GP XXI (period_int 21) and GP XXVI (period_int 26)
# Approximate: period 21 = FPÖ entered government 2000 but our data starts 2002;
# period 26 = 2017-19 coalition; year-based: 2002-2005 and 2017-2019

df_role <- df_fpoe |>
  filter(is_fpoe, !map_lgl(topics, is.null), !is.na(year)) |>
  mutate(
    role = case_when(
      year >= 2002 & year <= 2005 ~ "Government (2002–05)",
      year >= 2017 & year <= 2019 ~ "Government (2017–19)",
      TRUE ~ "Opposition"
    )
  ) |>
  unnest(topics) |>
  count(role, topics) |>
  group_by(role) |>
  mutate(pct = n / sum(n)) |>
  ungroup()

top_topics_role <- df_role |>
  summarise(total = sum(n), .by = topics) |>
  slice_max(total, n = 10, with_ties = FALSE) |>
  pull(topics)

df_role_top <- df_role |>
  filter(topics %in% top_topics_role) |>
  pivot_wider(
    id_cols = topics,
    names_from = role,
    values_from = pct,
    values_fill = 0
  ) |>
  mutate(
    oppo_minus_govt = Opposition - rowMeans(pick(starts_with("Government")))
  ) |>
  arrange(desc(abs(oppo_minus_govt)))

df_role_plot <- df_role |>
  filter(topics %in% top_topics_role) |>
  mutate(
    role_group = if_else(
      str_starts(role, "Government"),
      "Government",
      "Opposition"
    ),
    topics = factor(topics, levels = df_role_top$topics)
  ) |>
  summarise(pct = mean(pct), .by = c(topics, role_group))

ggplot(df_role_plot, aes(x = pct, y = topics, fill = role_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6, alpha = 0.85) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  scale_fill_manual(
    values = c("Government" = "#2171b5", "Opposition" = "#cb181d"),
    name = NULL
  ) +
  labs(
    x = "Share of FPÖ speeches on this topic",
    y = NULL,
    title = "In opposition, FPÖ talks more about migration and security",
    subtitle = "Top-10 topics by speech count; government vs. opposition periods averaged"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title.position = "plot",
    panel.grid.major.y = element_blank(),
    legend.position = "top"
  )
```

The contrast is stark. In opposition, FPÖ concentrates its speaking time on the topics most resonant with its voter base — migration, internal security, and European Union affairs. When in government, ministerial duties and coalition constraints push the party toward the full spectrum of policy, diluting the signature topics that define its brand. The data reflects a deliberate rhetorical strategy: maximise issue ownership pressure in opposition, manage policy breadth in government.

## Roll-call voting patterns

`get_item_details()` returns voting strings for items that received a plenary vote — recording which parties voted for and which voted against. This covers government bills (RV), parliamentary motions (ANTR), resolutions (E), and international treaty ratifications (RVS). Together these account for the full universe of contested legislative votes in the National Council.

```{r}
#| label: vote-data
#| cache: true
df_votes_raw <- readRDS("_cache_vote_rows_gp28.rds")
df_votes_long <- readRDS("_cache_votes_long_gp28.rds")

df_divided <- df_votes_long |>
  filter(vote_type == "divided", !is.na(vote))
```

### Party agreement

How often do two parties cast the same vote on contested items? The matrix below counts, for every pair, the share of divided votes where they chose identically.

```{r}
#| label: fig-vote-agreement
#| fig-cap: "Party vote agreement in the Austrian National Council, period XXVIII"
#| fig-height: 5
party_order <- c("ÖVP", "SPÖ", "FPÖ", "GRÜNE", "NEOS")

agreement_matrix <- df_divided |>
  select(item_url, party, vote) |>
  inner_join(
    df_divided |> select(item_url, party2 = party, vote2 = vote),
    by = "item_url",
    relationship = "many-to-many"
  ) |>
  filter(party != party2) |>
  group_by(party, party2) |>
  summarise(
    n_votes = n(),
    pct_agree = sum(vote == vote2) / n_votes,
    .groups = "drop"
  )

ggplot(
  agreement_matrix |>
    mutate(
      party = factor(party, levels = party_order),
      party2 = factor(party2, levels = rev(party_order))
    ),
  aes(x = party, y = party2, fill = pct_agree)
) +
  geom_tile(colour = "white", linewidth = 0.5) +
  geom_text(
    aes(label = scales::percent(pct_agree, accuracy = 1)),
    size = 3.5,
    colour = "white",
    fontface = "bold"
  ) +
  scale_fill_distiller(
    palette = "RdYlGn",
    direction = 1,
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1),
    name = "Agreement"
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Party vote agreement in the Austrian National Council",
    subtitle = sprintf(
      "Share of contested votes cast identically, NR period XXVIII (n=%d contested votes)",
      n_distinct(df_divided$item_url)
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title.position = "plot",
    axis.text = element_text(size = 11),
    legend.position = "right",
    panel.grid = element_blank()
  )
```

The coalition bond between ÖVP and SPÖ is near-perfect (99%), and NEOS votes with the governing parties on virtually every contested item. FPÖ agrees with other parties roughly half the time — which sounds moderate until you consider the composition: most of that "agreement" comes from FPÖ voting against opposition motions from GRÜNE and NEOS that the coalition also rejects. On government-initiated legislation, FPÖ is consistently on the opposite side.

### Who ends up on the winning side?

A more direct measure of parliamentary power: for each party, how often does their vote align with the final outcome?

```{r}
#| label: fig-vote-win-rate
#| fig-cap: "Win rate per party in contested votes, NR period XXVIII"
#| fig-height: 4
df_win_rate <- df_divided |>
  mutate(
    on_winning_side = (vote_result == "passed" & vote == "for") |
      (vote_result == "rejected" & vote == "against")
  ) |>
  group_by(party) |>
  summarise(
    n_votes = n(),
    pct_wins = sum(on_winning_side) / n_votes,
    .groups = "drop"
  ) |>
  arrange(pct_wins) |>
  mutate(party = factor(party, levels = party))

ggplot(df_win_rate, aes(x = pct_wins, y = party)) +
  geom_segment(
    aes(x = 0, xend = pct_wins, yend = party),
    colour = "grey75",
    linewidth = 1
  ) +
  geom_point(aes(colour = pct_wins), size = 5) +
  geom_text(
    aes(label = scales::percent(pct_wins, accuracy = 1)),
    hjust = -0.5,
    size = 3.5
  ) +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1.18),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_colour_distiller(
    palette = "RdYlGn",
    direction = 1,
    limits = c(0, 1),
    guide = "none"
  ) +
  coord_cartesian(clip = "off") +
  labs(
    x = "Share of contested votes on the winning side",
    y = NULL,
    title = "Who wins in contested votes?",
    subtitle = sprintf(
      "Share of divided votes where party voted with the majority, NR period XXVIII (n=%d)",
      n_distinct(df_divided$item_url)
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title.position = "plot", panel.grid.major.y = element_blank())
```

ÖVP, SPÖ, and NEOS win roughly two-thirds of contested votes — they form the effective parliamentary majority. GRÜNE wins about half; FPÖ wins fewer than one in three. The election winner of 2024 is, in voting terms, the most isolated party in parliament.

### How isolated is FPÖ?

```{r}
#| label: fig-vote-fpoe-isolation
#| fig-cap: "FPÖ isolation in contested votes, NR period XXVIII"
#| fig-height: 4
fpoe_votes <- df_divided |>
  filter(party == "FPÖ") |>
  select(item_url, fpoe_vote = vote)

fpoe_alignment <- fpoe_votes |>
  left_join(
    df_divided |> filter(party != "FPÖ") |> select(item_url, party, vote),
    by = "item_url"
  ) |>
  group_by(item_url, fpoe_vote) |>
  summarise(
    n_same = sum(vote == fpoe_vote, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    alignment = case_when(
      n_same == 4 ~ "All others agree",
      n_same >= 2 ~ "Majority agrees",
      n_same == 1 ~ "One party agrees",
      n_same == 0 ~ "FPÖ alone"
    )
  )

fpoe_alignment |>
  count(alignment) |>
  mutate(
    alignment = factor(
      alignment,
      levels = c(
        "FPÖ alone",
        "One party agrees",
        "Majority agrees",
        "All others agree"
      )
    ),
    pct = n / sum(n)
  ) |>
  ggplot(aes(x = pct, y = alignment, fill = alignment)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(
    aes(label = paste0(scales::percent(pct, accuracy = 1), " (n=", n, ")")),
    hjust = -0.1,
    size = 3.5
  ) +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.18))
  ) +
  scale_fill_manual(
    values = c(
      "All others agree" = "#2171b5",
      "Majority agrees" = "#6baed6",
      "One party agrees" = "#fdae6b",
      "FPÖ alone" = "#e6550d"
    )
  ) +
  labs(
    x = "Share of contested votes",
    y = NULL,
    title = "How isolated is FPÖ in its voting positions?",
    subtitle = "Contested votes in NR period XXVIII, by how many other parties voted with FPÖ"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title.position = "plot", panel.grid.major.y = element_blank())
```

In more than a third of contested votes, FPÖ stands completely alone — no other party casts the same vote. In a further 17%, only one other party agrees. Only when all four other parties happen to be on FPÖ's side (26% of cases) does FPÖ find itself in a genuine majority — and those are mostly the bipartisan items that transcend day-to-day politics.

### Where does parliament fight the most?

In period XXVIII, virtually every contested vote has at least one party in opposition — a sign of how polarised the chamber is. Rather than measuring which topics are *more* contested than others, the more informative question is where legislative energy is concentrated: which topics generate the most contested votes by volume?

```{r}
#| label: fig-vote-topics
#| fig-cap: "Topic areas with the most contested votes, NR period XXVIII"
#| fig-height: 6
df_topic_vol <- df_votes_raw |>
  filter(!map_lgl(topics, is.null), vote_type == "divided") |>
  unnest(topics) |>
  count(topics, name = "n_divided") |>
  slice_max(n_divided, n = 15) |>
  mutate(topics = forcats::fct_reorder(topics, n_divided))

ggplot(df_topic_vol, aes(x = n_divided, y = topics)) +
  geom_col(fill = "#cb181d", alpha = 0.8, width = 0.7) +
  geom_text(aes(label = n_divided), hjust = -0.3, size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(
    x = "Number of contested votes",
    y = NULL,
    title = "Where does parliament fight the most?",
    subtitle = "Topics with the most divided votes, NR period XXVIII"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title.position = "plot", panel.grid.major.y = element_blank())
```

Budget and finance dominate by sheer volume — the new coalition's fiscal agenda has generated more contested votes than any other area. Internal affairs, media, and climate/energy follow. Notably, many of these are areas where FPÖ has a strong ideological stake, and the high volume reflects both government-initiated legislation and FPÖ's counter-motions.

## Party names across history

Anyone working with Austrian parliamentary data spanning multiple decades will run into the naming problem. The FPÖ has been listed under various abbreviations over the years — "FPÖ", "F", "F-BZÖ" — depending on the period and context. The same applies to other parties.

The helper function `aux_parl_group_names_standard()` expands party abbreviations to include their historical variants. Several `get_*()` functions expose this via a `parl_group_names_standard` argument, which ensures that a search for "FPÖ" also captures "F" and "F-BZÖ", "NEOS" includes "NEOS-LIF", and so on. This matters more than you might think when aggregating data across legislative periods.

## Getting started

The [Get Started vignette](https://werkstattcodes.github.io/ParlAT/articles/ParlAT.html) walks through several worked examples — from counting questions by party across eight legislative periods to visualising government bills by topic, tracing MP careers through mandates, and exploring committee structures. The full function reference is available on the [documentation site](https://werkstattcodes.github.io/ParlAT/).

{ParlAT} is still a work in progress — the API surface of the Austrian Parliament is broad, documentation is sometimes sparse, and edge cases keep surfacing. That said, I hope the package is already a useful tool for anyone in the #rstats community working with Austrian parliamentary data, whether for research, journalism, teaching, or just curiosity about what happens in the Hohe Haus.

Feedback of any kind is very welcome: bug reports, questions, suggestions for new functions or better defaults, and pointers to use cases I haven't considered. The best place to reach me is the [GitHub issue tracker](https://github.com/werkstattcodes/ParlAT/issues) — opening an issue is the most reliable way to make sure nothing gets los - or contact me via direct message on [Bluesky](https://bsky.app/profile/zoowalk.bsky.social){target="_blank"}.