---
title: "Introducing {ParlAT}: An R Package to access Open Data of the Austrian Parliament"
date: 2026-02-14
date-modified: last-modified
author: Roland Schmidt
description: "{ParlAT} is an R package that wraps the Austrian Parliament's Open Data API — making it straightforward to query data on MPs, legislative items, committees, plenary sessions, and more from R."

categories:
 - Austria
 - parliament
 - r-bloggers

reference-location: margin
knitr:
  opts_chunk:
    code-fold: show
    collapse: true
    dev: "ragg_png"
draft: true
---

```{r}
#| label: setup
#| include: false
#| code-summary: "Setup"
library(ParlAT)
library(tidyverse)
```

Anyone who has followed this blog will know that I frequently rely on data published by the Austrian Parliament — whether it's for analysing [speeches](posts/2024-06-11-parliament-speech-length-limits/), tracing [calls to order](posts/2024-09-09-call-for-order/), or examining [the prior experience of MPs](posts/2024-10-24-mps-experience/). Until now, most of that work required custom scripts to query the Parliament's API and wrangle the JSON responses into tidy data frames. The [{ParlAT}](https://werkstattcodes.github.io/ParlAT/) package is an attempt to bundle these steps into a consistent, reusable interface. Overall, though, the package was developed with the intent to make the data of the Parliament more accessible to anyone interested in the Parliament's affairs, be it researchers, students, journalists etc, and working in R.

At this point, special thanks to Simon Hofer, Open Data evangelist at the Austrian Parliament, who was super helpful when diving into the details of the API. However, please note that neither the package nor its author is affiliated with the Austrian Parliament.

The package is still in development, but most of the functionality (and quite a bit more) is in place. Below I showcase **only some** examples.

You can install {ParlAT} from GitHub:

``` r
# install.packages("pak")
pak::pak("werkstattcodes/ParlAT")
```

For the full function reference and a more detailed introductory vignette, see the [documentation site](https://werkstattcodes.github.io/ParlAT/).

## What data is available?

The Austrian Parliament publishes [25 open datasets](https://www.parlament.gv.at/recherchieren/open-data/daten-und-lizenz/index.html) — with some exceptions — under a CC BY 4.0 license. These cover a broad range of parliamentary activity: MP records going back to 1918, committee compositions and memberships, plenary session protocols, citizen initiatives (Bürgerinitiativen), government bills (Regierungsvorlagen), written questions (Schriftliche Anfragen), EU submissions, and more.

{ParlAT} mirrors the API's structure. Most datasets map directly to a dedicated function; some, like the many types of parliamentary items (Verhandlungsgegenstände), are accessed through `get_items()` with different `item` parameters. Here's a condensed overview of the main functions:

| Function | Data |
|-----------------------------|-------------------------------------------|
| `get_mps()` | MPs since 1918, by period or date |
| `get_mps_current()` | Current composition of NR or BR |
| `get_mps_details()` | Individual MP speeches, activities, committee work |
| `get_mandates()` | All mandates of one or more individuals |
| `get_items()` | Parliamentary items: bills, motions, questions, etc. |
| `get_committees()` | Committees, optionally with member lists |
| `get_plenary_sessions()` | Plenary session metadata |
| `get_transcripts()` | Stenographic protocols with PDF export |
| `get_events()` | Parliamentary events and calendar |
| `get_participation()` | Public participation / consultation data |

## MPs: Current composition

A natural starting point is the composition of Parliament itself. `get_mps_current()` retrieves the members currently serving in the National Council (Nationalrat) or the Federal Council (Bundesrat):

```{r}
#| eval: true
#| code-summary: "Current NR composition"
df_current <- get_mps_current(institution = "NR")

glimpse(df_current)

df_current |>
  count(party, sort = TRUE)
```

The function also works for the Federal Council (Bundesrat):

```{r}
#| eval: true
#| code-summary: "Current BR composition"
df_current_br <- get_mps_current(institution = "BR")
nrow(df_current_br)
```

## MPs: Historical queries

For historical queries, `get_mps()` returns MPs active during a given legislative period or on a specific date. The function accepts both numeric and Roman numeral period identifiers (e.g., `27` or `"XXVII"`), as well as special codes like `"PN"` (Provisorische Nationalversammlung) and `"KN"` (Konstituierende Nationalversammlung).

```{r}
#| code-summary: "Composition of the National Council per date"
mpsEUMembership <- get_mps(date="10.06.1994", institution="NR")
nrow(mpsEUMembership)
glimpse(mpsEUMembership)
```

Unnesting the details column provides further info.

```{r}
#| code-summary: "Unnesting details"
mpsEUMembershipDetails <- mpsEUMembership %>%
  unnest(mp_details)

glimpse(mpsEUMembershipDetails)
```

### By legislative period

```{r}
#| eval: true
#| code-summary: "MPs by legislative period"
# All MPs of the National Council during the 22nd legislative period
mps_legis22 <- get_mps(legis_period = 22, institution = "NR")
```

Each row corresponds to one MP per legislative period. A nested `mp_details` column captures mandate-level details — party affiliation, electoral list, position — which can change within a single period. This is where things get interesting: unnesting `mp_details` reveals party switches, changes in parliamentary group, or mandate interruptions within a period.

### By date

When a specific date is more appropriate than a legislative period — for instance, when working with the Federal Council, which doesn't follow legislative periods in the same way — `get_mps()` accepts a `date` parameter:

```{r}
#| eval: true
#| code-summary: "MPs by date (BR)"
# Members of the Federal Council on 30 October 2013
df_mps_br <- get_mps(date = "30.10.2013", institution = "BR")
nrow(df_mps_br)

```

Next to the data retrieval functions, there are some auxiliary functions which hopefully come handy. E.g. to find the relevant dates for a specific legislative period, `get_legis_periods()` provides the start and end dates. If i have a date, it returns the pertaining legislative period.

```{r}
#| eval: true
#| code-summary: "Legislative period dates"
get_legis_periods(legis_period = 25)

get_legis_periods(date="29.03.1977")
```

### Filtering by gender, party, state

`get_mps()` supports a number of additional filters. These can be combined:

```{r}
#| eval: true
#| code-summary: "Filtering MPs by gender, party, state"
# Female SPÖ MPs in the current legislative period
female_spoe <- get_mps(
  gender = "female",
  party = "SPÖ",
  legis_period = 28,
  institution = "NR"
)
female_spoe

# MPs from Vienna during the 27th period
vienna_mps <- get_mps(
  state = "Wien",
  legis_period = 27,
  institution = "NR"
)
vienna_mps
```

## Mandates: Tracing careers across Parliament

`get_mandates()` returns all current and past mandates for one or more individuals. It accepts either names or unique `pad_intern` identifiers. Mandates are not limited to Parliament — they also cover executive positions (e.g., Chancellor, Minister), presidencies, and other functions.

```{r}
#| eval: true
#| code-summary: "Get mandates by name"
get_mandates(name = c("Karl Nehammer", "Andreas Babler"))
```

The returned data includes `position_code` (e.g., "NR", "BK" for Bundeskanzler, "BM" for Bundesminister), start and end dates, party affiliation, and a flag for whether the mandate is currently active. This makes it straightforward to calculate the total time an MP has served:

```{r}
#| eval: true
#| code-summary: "Service duration calculation"
# How long have current NR members served in the National Council?
df_mandates <- get_mandates(pad_intern = df_current$pad_intern, institution = "NR")

duration <- df_mandates |>
  # For active mandates, use today as end date
  mutate(position_date_end = case_when(
    is.na(position_date_end) & position_active == TRUE ~ today(),
    .default = position_date_end
  )) |>
  mutate(
    days = as.numeric(difftime(position_date_end, position_date_start, units = "days"))
  ) |>
  summarise(total_days = sum(days), .by = c(pad_intern, name)) |>
  arrange(desc(total_days))

# Top 5 longest-serving current NR members
duration |> slice_head(n = 5)
```

`get_mandates()` also accepts a `date` argument to filter for mandates active at a specific point in time, and an `institution` argument to restrict to a single chamber.

## Identifying MPs: `get_pad_intern()` and `get_names()`

Two further examples of auxiliary functions are `get_pat`

The unique identifier for each person in the Parliament's database is `pad_intern`. Many functions accept it as input. To look it up by name:

```{r}
#| eval: true
#| code-summary: "Look up pad_intern ID"
get_pad_intern("Strache")

```

`get_names()` retrieves all name variants an MP has used over time — relevant when names change due to marriage, divorce, or other reasons. A good example is Pia Philippa Beck (formerly Strache):

```{r}
#| eval: true
#| code-summary: "Name variants over time"
# All name variants

get_names(pad_intern = 44127)
```

## Parliamentary items (Verhandlungsgegenstände)

`get_items()` is arguably the most versatile function in the package. It covers a broad spectrum of parliamentary business — motions (Anträge), committee reports (Ausschussberichte), government bills (Regierungsvorlagen), written questions (Anfragen), EU submissions, citizen initiatives (Bürgerinitiativen), popular initiatives (Volksbegehren), and more — all controlled through the `item` parameter.

### Government bills

```{r}
#| eval: true
#| code-summary: "Government bills"
# Government bills tabled in the 27th legislative period
gov_bills <- get_items(item = "RV", legis_period = 27)

glimpse(gov_bills)
```

### Written questions across multiple periods

```{r}
#| eval: true
#| code-summary: "Written questions across periods"
# Written and oral questions from the 20th to 27th period
questions <- get_items(item = "J_JPR_M", legis_period = seq(20, 27))

nrow(questions)

```

The returned data includes metadata such as topic classification (`topics` is a list-column), sponsoring parliamentary group(s), and document type. This makes it straightforward to aggregate and visualise parliamentary activity — for example, counting questions by party and legislative period.

### Filtering by topic and parliamentary group

Items can be filtered by topic (Thema), parliamentary group, EuroVoc terms, and more. The `parl_group_names_standard` argument expands party abbreviations to their historical variants — so searching for "FPÖ" also captures items filed under "F" or "F-BZÖ":

```{r}
#| eval: true
#| code-summary: "Filter by topic and party"
# SPÖ motions on health topics during the 27th period
spoe_health <- get_items(
  item = "ANTR",
  topic = "Gesundheit und Ernährung",
  parl_group = "SPÖ",
  legis_period = 27,
  institution = "NR",
  parl_group_names_standard = TRUE
)
```

## An MP's work in detail

`get_mps_details()` provides granular data on individual MPs' plenary speeches, committee memberships, and parliamentary activities. It takes a `pad_intern` and a `detail_type` ("plenary", "committees", or "activities").

### Plenary speeches

```{r}
#| eval: true
#| code-summary: "Plenary speeches"
# All plenary speeches by Sebastian Kurz in the National Council
get_pad_intern("Krisper")

df_plenary_krisper <- get_mps_details(
  pad_intern = 2344,
  detail_type = "plenary",
  institution = "NR"
)

# Kurz held speeches in different capacities — Abgeordneter, Bundeskanzler, etc.
df_plenary_krisper |>
  unnest(position_name) |>
  count(position_name, sort = TRUE)
```

`get_mps_details()` returns all speeches regardless of the person's role at the time. The `position_name` list-column reveals which position(s) the person held when the speech was given.

### Parliamentary activities

```{r}
#| eval: true
#| code-summary: "Parliamentary activities"
# Written questions submitted by an MP

get_pad_intern("Pilz")

df_activities <- get_mps_details(
  pad_intern = 1210,
  detail_type = "activities",
  institution = "NR",
  item = "JMIN" # Written questions to the Federal Government
)

nrow(df_activities)

glimpse(df_activities)
```

## Committees (Ausschüsse)

`get_committees()` retrieves committee information for a given legislative period and institution. It can also extract membership lists when `details_type = "members"` is set.

### Listing committees

```{r}
#| eval: true
#| code-summary: "List committees"
# All committees of the National Council in the 27th period
committees_nr27 <- get_committees(legis_period = 27, institution = "NR")
committees_nr27
```

### Committees of inquiry (Untersuchungsausschüsse)

The `citation` parameter allows filtering by committee type code. Committees of inquiry use the code "USA":

```{r}
#| eval: true
#| code-summary: "Committees of inquiry"
# How many committees of inquiry per legislative period?
inquiry_committees <- map(
  seq(20, 28),
  \(x) get_committees(legis_period = x, institution = "NR", citation = "USA")
) |>
  list_rbind()

inquiry_committees |>
  mutate(legis_period = factor(legis_period, levels = as.character(as.roman(20:28)))) |>
  count(legis_period, .drop = FALSE)
```

## Plenary sessions

`get_plenary_sessions()` retrieves metadata about plenary sessions — dates, session numbers, and links to agendas. It supports three modes via the `session_and_activities` parameter:

-   `"sessions"`: Basic session metadata (dates, numbers, agenda URLs)
-   `"submitted"`: Items submitted during sessions (motions, questions, etc.)
-   `"held"`: Types of proceedings held (Question Hours, Government Declarations, etc.)

```{r}
#| eval: true
#| code-summary: "Plenary sessions"
# Plenary sessions from the 20th to 27th period

sessions <- get_plenary_sessions(
  institution = "NR",
  legis_period = seq(20, 27),
  session_and_activities = "sessions"
)

# Count sessions per legislative period
sessions |> count(legis_period)

# What types of motions were submitted during 27th period sessions?
submitted_27 <- get_plenary_sessions(
  institution = "NR",
  legis_period = 27,
  session_and_activities = "submitted",
  submitted = "UEA" # Dependent resolution motions
)
glimpse(submitted_27)
```

## Transcripts (Stenographische Protokolle)

`get_transcripts()` provides access to stenographic protocols. Results include links to both HTML and PDF versions of the transcripts. For bulk downloads, the function supports PDF export

```{r}
#| eval: false
#| code-summary: "Search transcripts"
# Transcripts mentioning "Gesundheit" in the 28th period
transcripts <- get_transcripts(
  search_string = "gesundheit",
  legis_period = 28,
  session_type = "NRSITZ",
  echo = TRUE
)

glimpse(transcripts)
```

When defining a destination folder, the function provides for the download of the transcript files.

```{r}
#| eval: false
#| code-summary: "Download transcript PDFs"

# Download PDFs to a local folder
get_transcripts(
  legis_period = 28,
  session_type = "NRSITZ",
  export = "pdf",
  export_destination = "my_transcripts"
)
```

## The `echo` argument

Most functions accept an `echo` argument. When set to `TRUE`, the function prints the search parameters and a direct URL linking to the same results on the Parliament's website:

```{r}
#| eval: false
#| code-summary: "Echo parameter example"
get_items(item = "RV", legis_period = 27, echo = TRUE)
#> {"NRBR":"NR","GP":"XXVII.","VHG":"RV"}
#> https://www.parlament.gv.at/recherchieren/gegenstaende/...
#> [1] 245
```

I find this useful for verification, for quickly sharing a specific query with someone who prefers the web interface, and for sanity-checking that the API query is doing what I expect. It's a small feature but it's saved me more than once.

## Party names across history

Anyone working with Austrian parliamentary data spanning multiple decades will run into the naming problem. The FPÖ has been listed under various abbreviations over the years — "FPÖ", "F", "F-BZÖ" — depending on the period and context. The same applies to other parties.

The helper function `aux_parl_group_names_standard()` expands party abbreviations to include their historical variants. Several `get_*()` functions expose this via a `parl_group_names_standard` argument, which ensures that a search for "FPÖ" also captures "F" and "F-BZÖ", "NEOS" includes "NEOS-LIF", and so on. This matters more than you might think when aggregating data across legislative periods.

## Getting started

The [Get Started vignette](https://werkstattcodes.github.io/ParlAT/articles/ParlAT.html) walks through several worked examples — from counting questions by party across eight legislative periods to visualising government bills by topic, tracing MP careers through mandates, and exploring committee structures. The full function reference is available on the [documentation site](https://werkstattcodes.github.io/ParlAT/).

{ParlAT} is still in active development. If you encounter issues or have ideas for additional features, feel free to open an issue on [GitHub](https://github.com/werkstattcodes/ParlAT/issues).